---
title: "Readr solution"
author: "Suganya"
date: '2022-09-16'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


What function would you use to read a file where fields are separated with “|”?

```{r}
read_delim("a|b|c\n1|2|3\n4|5|6", delim = "|")
```



Apart from file, skip, and comment, what other arguments do read_csv() and read_tsv() have in common?

read_csv() and read_tsv() have the same arguments. They only difference is that one is comma delimited, and the 

other is tab delimited.

3 – What are the most important arguments to read_fwf()?

The most important argument is col_positions, which defines the column positions.

4 – Sometimes strings in a CSV file contain commas. To prevent them from causing problems they need to be surrounded by a quoting character, like " or '. By convention, read_csv() assumes that the quoting character will be ", and if you want to change it you’ll need to use read_delim() instead. What arguments do you need to specify to read the following text into a data frame?

```{r}
read_csv("x,y\n1,'a,b'", quote = "'")
```

5 – Identify what is wrong with each of the following inline CSV files. What happens when you run the code?

read_csv("a,b\n1,2,3\n4,5,6")

Only two columns names are provided, so the values in the last column are dropped.

read_csv("a,b,c\n1,2\n1,2,3,4")

Only three column names are provided. The value in the last column in the last row is dropped, and NA is coerced in the third column of second row.

read_csv("a,b\n\"1")

The open quote \" is dropped because there is no paired close quote. There is only one value in the second row, so NA is coerced in the second column

read_csv("a,b\n1,2\na,b")

Since the second rows are strings, the entire columns are coerced into strings.

read_csv("a;b\n1;3")

read_csv() looks for commas, not semi-colons. Everything is treated as one column name and one value.






1. What are the most important arguments to locale()?

The locale object has arguments to set the following:

date and time formats: date_names, date_format, and time_format
time zone: tz
numbers: decimal_mark, grouping_mark
encoding: encoding

2. What happens if you try and set decimal_mark and grouping_mark to the same character? What happens to the default
value of grouping_mark when you set decimal_mark to ",“?

What happens to the default value of decimal_mark when you set the grouping_mark to ".“?

If the decimal and grouping marks are set to the same character, locale throws an error:

```{r}
locale(decimal_mark = ".", grouping_mark = ".")
```

I didn’t discuss the date_format and time_format options to locale(). What do they do? Construct an example that shows when they might be useful.

```{r}
locale_custom <- locale(date_format = "Day %d Mon %M Year %y",
                 time_format = "Sec %S Min %M Hour %H")
date_custom <- c("Day 01 Mon 02 Year 03", "Day 03 Mon 01 Year 01")
parse_date(date_custom)
```
```{r}
parse_date(date_custom, locale = locale_custom)
```
```{r}
time_custom <- c("Sec 01 Min 02 Hour 03", "Sec 03 Min 02 Hour 01")
parse_time(time_custom)
```
```{r}
parse_time(time_custom, locale = locale_custom)
```

4. If you live outside the US, create a new locale object that encap‐
sulates the settings for the types of files you read most com‐
monly

As an example, consider Australia. Most of the defaults values are valid, except that the date format is “(d)d/mm/yyyy”, meaning that January 2, 2006 is written as 02/01/2006.
```{r}
parse_date("02/01/2006")
```

To correctly parse Australian dates, define a new locale object.

```{r}
au_locale <- locale(date_format = "%d/%m/%Y")
parse_date("02/01/2006", locale = au_locale)
```

What’s the difference between read_csv() and read_csv2()?

The delimiter. The function read_csv() uses a comma, while read_csv2() uses a semi-colon (;). Using a semi-colon is useful when commas are used as the decimal point (as in Europe).


What are the most common encodings used in Europe? What are the most common encodings used in Asia? Do some googling to find out.

UTF-8 is standard now, and ASCII has been around forever.

For the European languages, there are separate encodings for Romance languages and Eastern European languages using Latin script, Cyrillic, Greek, Hebrew, Turkish: usually with separate ISO and Windows encoding standards. There is also Mac OS Roman.

For Asian languages Arabic and Vietnamese have ISO and Windows standards




Generate the correct format string to parse each of the following dates and times:




```{r}
d1 <- "January 1, 2010"
d2 <- "2015-Mar-07"
d3 <- "06-Jun-2017"
d4 <- c("August 19 (2015)", "July 1 (2015)")
d5 <- "12/30/14" # Dec 30, 2014
t1 <- "1705"
t2 <- "11:15:10.12 PM"
```

```{r}
parse_date(d1, "%B %d, %Y")

parse_date(d2, "%Y-%b-%d")

parse_date(d3, "%d-%b-%Y")

parse_date(d4, "%B %d (%Y)")

parse_date(d5, "%m/%d/%y")

parse_time(t1, "%H%M")
```

