  ---
title: "Data Import with readr"
author: "Sathish G"
date: "19/11/2019"
output: html_document
---

# Data Import with readr

## Introduction

- This chapter deals with how to read plain-text rectangular files into R.

### Prerequisites

```{r include=FALSE}
library(tidyverse)
```


### Getting Started

- read_csv() - reads comma-delimited files
  read_csv2() - reads semicolon separated files
  read_tsv() - reads tab-delimited files
  read_delim() reads in files with any delimiter

- read_fwf() - reads fixed-width files

- read_log() - reads Apache style log files

Note: These functions will have similar syntax. Once you've mastered one, you can use the others with ease.

**read_csv()**:
- first argument -> the path to the file to read.
- Example:

heights <- read_csv("data/heights.csv")

```{r}
#heights <- read_csv("/home/sathish/Desktop/heights.csv")
```

```{r}
#heights
```


**Inline CSV file**:

```{r}
read_csv("a,b,c
         1,2,3
         4,5,6")
```

In both the cases read_csv() uses the first line of the data for the column names.

Two cases, where we might want to change this behavior:

1. Sometimes there are a few lines of metadata at the top of the file.

```{r}
read_csv("The first line of metadata
         The second line of metadata
         x,y,z
         1,2,3", skip = 2)
```

```{r}
read_csv("# A comment I want to skip
         x,y,z
         1,2,3", comment = "#")
```

2. The data might not have column names. You can use col_names = FALSE to tell read_csv() not to treat the first row as headings and instead label them sequentially from X1 to Xn

```{r}
read_csv("1,2,3\n4,5,6", col_names = FALSE)
```

Alternatively, you can pass col_names a character vector, which will be used as the column names.

```{r}
read_csv("1,2,3\n4,5,6", col_names = c("x","y","z"))
```

```{r}
read_csv("a,b,c\n1,2,.",na = ".")
```


### Compared to Base R

There are few good reasons to favor readr functions over the base equivalents like read.csv():

- They are typically much faster(~10x) than their base equivalents.
  Long running jobs have a progress bar, so you can see what's happening.

- They produce tibbles, and they don't convert character vectors to factors, use row names, or munge the column names.

- They are more reproducible. Base R functions inherit some behavior from your operating system and environment variables, so import code that works on your computer might not work on someone else's.

## Parsing a Vector

parse_*() functions - take a character vector and return a more specialized vector like a logical, integer or date

```{r}
str(parse_logical(c("TRUE","FALSE","NA")))
```

```{r}
str(parse_integer(c("1","2","3")))
```

```{r}
str(parse_date(c("2010-01-01","1979-10-14")))
```

- the first argument is a character vector to parse
- na argument specifies which strings should be treated as missing

```{r}
parse_integer(c("1","231",".","456"), na = ".")
```

- if parsing fails, you'll get a warning:

```{r}
x <- parse_integer(c("123","345","abc","123.45"))
```

- And the failures will be missing in the output

```{r}
x
```

```{r}
problems(x)
```

There are 8 particularly important parsers:

1. parse_logical() and parse_integer()
2. parse_double() and parse_number()
3. parse_character()
4. parse_factor()
5. parse_datetime(), parse_date(), parse_time()

### Numbers

1. People write numbers differently in different parts of the world (. and ,)

```{r}
parse_double("1.23")
```
```{r}
parse_double("1,23", locale = locale(decimal_mark = ","))
```

2. Numbers are often surrounded by other characters that provide some context, like "$1000" or "10%"

```{r}
parse_number("$100")
```

```{r}
parse_number("20%")
```

```{r}
parse_number("It cost $123.45")
```

3. Numbers often contain "grouping" characters to make them easier to read, like "1,000,000"

```{r}
# used in america
parse_number("$123,456,789")
```

```{r}
# used in many parts of europe
parse_number("123.456.789", locale = locale(grouping_mark = "."))
```

```{r}
# used in switzerland
parse_number("123'456'789", locale = locale(grouping_mark = "'"))
```


### Strings

```{r}
charToRaw("Hadley")
```

```{r}
x1 <- "El Ni\xf1o was particularly bad this year"
x2 <- "\x82\xb1\x82\xf1\x82\xc9\x82\xbf\x82\xcd"
```

```{r}
parse_character(x1, locale = locale(encoding = "Latin1"))
parse_character(x2, locale = locale(encoding = "Shift-JIS"))
```

```{r}
guess_encoding(charToRaw(x1))
```

```{r}
guess_encoding(charToRaw(x2))
```


### Factors

R uses factors to represent categorical variables that have a known set of possible values.
Give parse_vector() a vector of known levels to generate a warning whenever an unexpected value is present:

```{r}
fruit <- c("apple", "banana")
parse_factor(c("apple","banana","bananana"), levels = fruit)
```

### Parsing a File

```{r}
guess_parser("2010-10-01")
guess_parser("15:01")
guess_parser(c("TRUE","FALSE"))
guess_parser(c("1","5","9"))
guess_parser(c("12,352,561"))
```

```{r}
str(parse_guess("2010-10-10"))
```



```{r}
#setwd("C:/Users/suganya/Desktop/R Programming")
file1 <- read_csv("insurance.csv")
#file1 <- read_csv(readr_example("insurance.csv"))
problems(file1)
tail(file1)
```
```{r}
setwd("C:/Users/suganya/Desktop/R Programming")
challenge2 <- read_csv(
 "insurance.csv",
 guess_max = 1001
 )


```
```{r}
write_csv(file1, "challenge.csv")

```

### Writing to a File

write_csv(challenge, "challenge.csv")

first argument - the data frame to save
second argument - the location or path to save its