---
title: "Tibble and Solutions"
author: "Suganya"
date: '2022-09-16'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Tibbles with tibble

```{r}
library(tidyverse)
```

Creating Tibbles
```{r}
as_tibble(iris)
```
```{r}
tibble(
 x = 1:5,
 y = 1,
 z = x ^ 2 + y
)

```

```{r}
tb <- tibble(
 `:)` = "smile",
 ` ` = "space",
 `2000` = "number"
)
tb
```
Tibbles Versus data.frame

There are two main differences in the usage of a tibble versus a clas‐
sic data.frame: printing and subsetting.

Printing
Tibbles have a refined print method that shows only the first 10
rows, and all the columns that fit on screen

```{r}
tibble(
 a = lubridate::now() + runif(1e3) * 86400,
 b = lubridate::today() + runif(1e3) * 30,
 c = 1:1e3,
 d = runif(1e3),
 e = sample(letters, 1e3, replace = TRUE)
)
```

Subsetting

If you want to pull out a single variable, you need some new
tools, $ and [[. [[ can extract by name or position; $ only extracts
by name but is a little less typing:

```{r}
df <- tibble(
 x = runif(5),
 y = rnorm(5)
)

```

```{r}
df$x
df[["x"]]
df[[1]]

```


Exercises

1. How can you tell if an object is a tibble? (Hint: try printing
mtcars, which is a regular data frame.)

```{r}
class(mtcars)
```


2 – Compare and contrast the following operations on a data.frame and equivalent tibble. What is different? Why might the default data frame behaviours cause you frustration?

Creating tibble in R is pretty much the same as creating a data.frame object.

```{r}
df <- data.frame(abc = 1, xyz = "a")
tdf <- tibble(abc = 1, xyz = "a")
```

```{r}
str(df)
str(tdf)
```
If you have the name of a variable stored in an object, e.g., var <- "mpg", how can you extract the reference variable from a tibble?

```{r}
tibble_mtcars <- as.tibble(mtcars)
var <- 'mpg'
tibble_mtcars[var]
```


Practice referring to nonsyntactic names in the following data
frame by:
a. Extracting the variable called 1.
b. Plotting a scatterplot of 1 versus 2.
c. Creating a new column called 3, which is 2 divided by 1.
d. Renaming the columns to one, two, and three:
annoying <- tibble(
 `1` = 1:10,
 `2` = `1` * 2 + rnorm(length(`1`))
)

```{r}
annoying <- tibble(
  `1` = 1:10,
  `2` = `1` * 2 + rnorm(length(`1`))
)
```

```{r}
annoying$`1`
```

```{r}
annoying %>% ggplot() +
  geom_point(mapping = aes(x = `1`, y = `2`))
```

```{r}
annoying$`3` <- annoying$`2` / annoying$`1`
annoying
```

```{r}
annoying %>% rename(one = `1`, two = `2`, three = `3`)
```


What does tibble::enframe() do? When might you use it?

```{r}
x <- c(Joe = 24, May = 33, Jack = 55)
enframe(x, name = 'Name', value = 'Age')
```
6 – What option controls how many additional column names are printed at the footer of a tibble?

By default, information of all remaining columns are printed at the footer. To limit the number of additional column information, we can use the argument n_extra. 

```{r}
print(nycflights13::flights, n_extra = 2)
```

What function would you use to read a file where fields are separated with “|”?

```{r}
read_delim("a|b|c\n1|2|3\n4|5|6", delim = "|")
```



Apart from file, skip, and comment, what other arguments do read_csv() and read_tsv() have in common?

read_csv() and read_tsv() have the same arguments. They only difference is that one is comma delimited, and the 

other is tab delimited.

3 – What are the most important arguments to read_fwf()?

The most important argument is col_positions, which defines the column positions.

4 – Sometimes strings in a CSV file contain commas. To prevent them from causing problems they need to be surrounded by a quoting character, like " or '. By convention, read_csv() assumes that the quoting character will be ", and if you want to change it you’ll need to use read_delim() instead. What arguments do you need to specify to read the following text into a data frame?

```{r}
read_csv("x,y\n1,'a,b'", quote = "'")
```

5 – Identify what is wrong with each of the following inline CSV files. What happens when you run the code?

read_csv("a,b\n1,2,3\n4,5,6")

Only two columns names are provided, so the values in the last column are dropped.

read_csv("a,b,c\n1,2\n1,2,3,4")

Only three column names are provided. The value in the last column in the last row is dropped, and NA is coerced in the third column of second row.

read_csv("a,b\n\"1")

The open quote \" is dropped because there is no paired close quote. There is only one value in the second row, so NA is coerced in the second column

read_csv("a,b\n1,2\na,b")

Since the second rows are strings, the entire columns are coerced into strings.

read_csv("a;b\n1;3")

read_csv() looks for commas, not semi-colons. Everything is treated as one column name and one value.


