---
title: "Unit III-Solution"
author: "Suganya"
date: '2022-10-01'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


1.How can you tell if an object is a tibble? (Hint: try printing mtcars, which is a regular data frame).

When  print mtcars, it prints all the columns

But when we first convert mtcars to a tibble using as_tibble(), 

it prints on the first ten observations. There are also some other differences in formatting of the printed data frame.

```{r}


```

Use the function is_tibble() to check whether a data frame is a tibble or not. 

The mtcars data frame is not a tibble.

```{r}

```

But the diamonds and flights data are tibbles.

```{r}
is_tibble(ggplot2::diamonds) 
```


```{r}
is_tibble(nycflights13::flights) 
```
 

```{r}
is_tibble(as_tibble(mtcars)) 
```
 

More generally, you can use the class() function to find out the class of an object.

```{r}
class(mtcars) 
```

2. Compare and contrast the following operations on a data.frame and equivalent tibble. What is different? Why might

the default data frame behaviors cause you frustration?

```{r}
df <- data.frame(abc = 1, xyz = "a") 
df$x 
df[, "xyz"] 
df[, c("abc", "xyz")]
```



```{r}
df <- data.frame(abc = 1, xyz = "a") 
df=as_tibble(df)
df$x 
df[, "xyz"] 
df[, c("abc", "xyz")]
```

3. If you have the name of a variable stored in an object, e.g. var <- "mpg", how can you extract the reference

variable from a tibble?

Use the double bracket, like var[[1]].

Cannot use the dollar sign, because var$mpg would look for a column named mpg.

4. Practice referring to non-syntactic names in the following data frame by:

A. Extracting the variable called 1.

For this example, create a dataset called exampe with columns named 1 and 2.

```{r}

```


To extract the variable named 1:

```{r}

```


B. Plotting a scatterplot of 1 vs 2.

```{r}

```


Creating a new column called 3 which is 2 divided by 1.

```{r}

```


Renaming the columns to one, two and three.

```{r}

```


5. What does tibble::enframe() do? When might you use it?

The function tibble::enframe() converts named vectors to a data frame with names and values

```{r}

```

6. What option controls how many additional column names are printed at the footer of a tibble?

The help page for the print() method of tibble objects is discussed in ?print.tbl. The n_extra argument determines

the number of extra columns to print information for.
```{r}
library(nycflights13)
```

```{r}

```


1. What function would you use to read a file where fields are separated with “|”?

```{r}

```

2. Apart from file, skip, and comment, what other arguments do read_csv() and read_tsv() have in common?

Execute the command to know the remaining arguments:

```{r}
union(names(formals(read_csv)), names(formals(read_tsv))) 
```
col_names and col_types are used to specify the column names and how to parse the columns

locale is important for determining things like the encoding and whether “.” or “,” is used as a decimal mark.

na and quoted_na control which strings are treated as missing values when parsing vectors

trim_ws trims whitespace before and after cells before parsing

n_max sets how many rows to read

guess_max sets how many rows to use when guessing the column type

progress determines whether a progress bar is shown

3. What are the most important arguments to read_fwf()?

The most important argument to read_fwf() which reads “fixed-width formats”, is col_positions which tells the

function where data columns begin and end.

4. Sometimes strings in a CSV file contain commas. To prevent them from causing problems they need to be surrounded

by a quoting character, like " or '. By convention, read_csv() assumes that the quoting character will be ", and if

you want to change it you’ll need to use read_delim() instead. What arguments do you need to specify to read the

following text into a data frame?

 "x,y\n1,'a,b'“
 
For read_delim(), need to specify a delimiter, in this case ",", and a quote argument.

```{r}


```
 
5. Identify what is wrong with each of the following inline CSV files.

 What happens when you run the code?
 
```{r}
read_csv("a,b\n1,2,3\n4,5,6") 
```
 
 


```{r}
read_csv("a,b,c\n1,2\n1,2,3,4") 
```



```{r}
read_csv("a,b\n\"1") 
```


```{r}
read_csv("a,b\n1,2\na,b")


```

```{r}
read_csv("a;b\n1;3")

```



1. What are the most important arguments to locale()?

The locale object has arguments to set the following:

date and time formats: date_names, date_format, and time_format
time zone: tz
numbers: decimal_mark, grouping_mark
encoding: encoding



2. What happens if you try and set decimal_mark and grouping_mark to the same character? What happens to the default
value of grouping_mark when you set decimal_mark to ",“?

What happens to the default value of decimal_mark when you set the grouping_mark to ".“?

If the decimal and grouping marks are set to the same character, locale throws an error:



```{r}
locale(decimal_mark = ".", grouping_mark = ".")
```

I didn’t discuss the date_format and time_format options to locale(). What do they do? Construct an example that shows when they might be useful.

```{r}
locale_custom <- locale(date_format = "Day %d Mon %M Year %y",
                 time_format = "Sec %S Min %M Hour %H")
date_custom <- c("Day 01 Mon 02 Year 03", "Day 03 Mon 01 Year 01")
parse_date(date_custom)
```
```{r}
parse_date(date_custom, locale = locale_custom)
```
```{r}
time_custom <- c("Sec 01 Min 02 Hour 03", "Sec 03 Min 02 Hour 01")
parse_time(time_custom)
```
```{r}
parse_time(time_custom, locale = locale_custom)
```

4. If you live outside the US, create a new locale object that encap‐
sulates the settings for the types of files you read most com‐
monly

As an example, consider Australia. Most of the defaults values are valid, except that the date format is “(d)d/mm/yyyy”, meaning that January 2, 2006 is written as 02/01/2006.
```{r}
parse_date("02/01/2006")
```

To correctly parse Australian dates, define a new locale object.

```{r}
au_locale <- locale(date_format = "%d/%m/%Y")
parse_date("02/01/2006", locale = au_locale)
```

What’s the difference between read_csv() and read_csv2()?

The delimiter. The function read_csv() uses a comma, while read_csv2() uses a semi-colon (;). Using a semi-colon is useful when commas are used as the decimal point (as in Europe).


What are the most common encodings used in Europe? What are the most common encodings used in Asia? Do some googling to find out.

UTF-8 is standard now, and ASCII has been around forever.

For the European languages, there are separate encodings for Romance languages and Eastern European languages using Latin script, Cyrillic, Greek, Hebrew, Turkish: usually with separate ISO and Windows encoding standards. There is also Mac OS Roman.

For Asian languages Arabic and Vietnamese have ISO and Windows standards




Generate the correct format string to parse each of the following dates and times:




```{r}
d1 <- "January 1, 2010"
d2 <- "2015-Mar-07"
d3 <- "06-Jun-2017"
d4 <- c("August 19 (2015)", "July 1 (2015)")
d5 <- "12/30/14" # Dec 30, 2014
t1 <- "1705"
t2 <- "11:15:10.12 PM"
```