---
title: "Untitled"
author: "Suganya"
date: '2022-10-01'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

A data frame is simply a matrix, but can have columns with different types (numeric, character, logical). 

Rows are observations (individuals) and columns are variables. 

A tibble is a subtype of a data frame that is optimized for data science applications

Tibbles have a refined print method that shows only the first 10 rows, and all the columns that fit on screen.

This makes it much easier to work with large data. 

```{r}
library(tidyverse) 
#install.packages("tibble")
library(tibble)

```

Force a data frame to a tibble
```{r}
as_tibble(iris)

```
```{r}
tibble( x = 1:5, y = 1, z = x ^ 2 + y ) 

```

It’s possible for a tibble to have column names that are not valid R variable names
```{r}
tb <- tibble( `:)` = "smile", ` ` = "space", `2000` = "number" )
```

There are two main differences in the usage of a tibble versus a classic data.frame: printing and subsetting.

Printing :Tibbles have a refined print method that shows only the first 10 rows, and all the columns that fit on

screen.

```{r}
as_tibble(iris)
```

Subsetting :If you want to pull out a single variable, you need some new tools, $ and [[. 

[[ can extract by name or position; 

$ only extracts by name but is a little less typing

Compared to a data.frame, tibbles are more strict: 

they never do partial matching, and they will generate a warning if the column to access does not exist.

```{r}
tb$`2000`
tb[["2000"]]
```
Some older functions don’t work with tibbles. 

Use as.data.frame() to turn a tibble back to a data.frame

```{r}
class(as.data.frame(tb))
```
 
 This chapter concern with how to read plain-text rectangular files into R


```{r}
library(tidyverse)
#install.packages(“readr”)
library(readr)

```

Most of readr’s functions are concerned with turning flat files into data frames

readr’s functions 

read_csv() reads comma-delimited files
read_csv2() reads semicolon-separated files 
read_tsv() reads tab-delimited files
read_delim() reads in files with any delimiter

read_fwf() reads fixed-width files. 

Specify fields either by their widths with fwf_widths() or their position with fwf_positions(). 

read_table() reads a common variation of fixed-width files where columns are separated by white space. 

read_log() reads Apache style log files. 

The first argument to read_csv() is the path to the file to read:

heights <- read_csv("data/heights.csv") 

Prints out a column specification that gives the name and type of each column

Inline CSV file--for creating reproducible examples

```{r}
read_csv("a,b,c 
				1,2,3 
				4,5,6")

```

skip = n to skip the first n lines; 

or use comment = "#" to drop all lines that start with (e.g.) #:

```{r}
read_csv("The first line of metadata 
				 The second line of metadata  
				x,y,z  
				1,2,3", skip = 2)
```

```{r}
read_csv("# A comment I want to skip  
				x,y,z  
				1,2,3", comment = "#")

```

Use col_names = FALSE to tell read_csv() not to treat the first row as headings, and instead label them sequentially

from X1 to Xn: 

```{r}
read_csv("1,2,3\n4,5,6", col_names = FALSE)
```


Pass col_names a character vector,  which will be used as the column names:

```{r}
 read_csv("1,2,3\n4,5,6", col_names = c("x", "y", "z"))
```

To represent missing values in file:

```{r}
read_csv("a,b,c\n1,2,.", na = ".")
```

There are a few good reasons to favor readr functions over the base equivalents: 

They are typically much faster (~10x) than their base equivalents.

They produce tibbles, and they don’t convert character vectors to factors, use row names, or munge the column names.

These are common sources of frustration with the base R functions.

They are more reproducible. Base R functions inherit some behavior from your operating system and environment

variables, so import code that works on your computer might not work on someone else’s.


## Parsing a Vector

These functions take a character vector and return a more specialized vector like a logical, integer, or date:

```{r}
str(parse_logical(c("TRUE", "FALSE", "NA"))) 

```

```{r}
str(parse_integer(c("1", "2", "3")))
```

 
```{r}
 str(parse_date(c("2010-01-01", "1979-10-14"))) 
```


These functions take a character vector and return a more specialized vector like a logical, integer, or date:

```{r}
str(parse_logical(c("TRUE", "FALSE", "NA"))) 
```



```{r}
str(parse_integer(c("1", "2", "3")))
```

 
 
```{r}
str(parse_date(c("2010-01-01", "1979-10-14"))) 
```
 
na argument specifies which strings should be treated as missing:

```{r}
parse_integer(c("1", "231", ".", "456"), na = ".") 
```

If parsing fails, get a warning &  the failures will be missing in the output

```{r}
x <- parse_integer(c("123", "345", "abc", "123.45")) 
```

If there are many parsing failures, use problems() to get the complete set. 

This returns a tibble, which you can then manipulate with dplyr:
```{r}
problems(x) 

```

There are eight particularly important parsers:

parse_logical() and parse_integer() parse logicals and integers, respectively. 

parse_double() is a strict numeric parser, and parse_number() is a flexible numeric parser. 

parse_character()  to parse characters

parse_factor() creates factors

parse_datetime(), parse_date(), and parse_time() allows to parse various date and time specifications.  



Three problems to parse a number, 

People write numbers differently in different parts of the world. For example, some countries use . in between the

integer and fractional parts of a real number, while others use ,. 

Numbers are often surrounded by other characters that provide some context, like “$1000” or “10%”.


Numbers often contain “grouping” characters to make them easier to read, like “1,000,000”, and these grouping

characters vary around the world. 

To address the first problem, readr has the notion of a “locale,” and  setting the decimal_mark argument:

```{r}
parse_double("1.23") 

```

```{r}
parse_double("1,23", locale = locale(decimal_mark = ",")) 
```


parse_number() addresses the second problem: it ignores nonnumeric characters before and after the number

```{r}
parse_number("$100")
 #> [1] 100
 parse_number("20%")
 #> [1] 20
 parse_number("It cost $123.45")
 #> [1] 123
```


The final problem is addressed by the combination of parse_num ber() and the locale as parse_number() will ignore

the “grouping mark”:

```{r}
# Used in America 
parse_number("$123,456,789") 
#> [1] 1.23e+08
# Used in many parts of Europe
 parse_number(  "123.456.789",  locale = locale(grouping_mark = ".") ) 
#> [1] 1.23e+08
# Used in Switzerland 
parse_number(  "123'456'789",  locale = locale(grouping_mark = "'") )
 #> [1] 1.23e+08


```


