---
title: "Tidy Data with tidyr"
author: "Sathish G"
date: "20/09/2019"
output: html_document
---

# Tidy Data with tidyr

> Tidy datasets are all alike, but every messy dataset is messy in its own way.
> - Hadley Wickham

**Attaching the Package**

```{r include=FALSE}
library(tidyr)
library(tidyverse)
```

## Tidy Data

- Each dataset shows the same values of four variables, country, year, population, and cases, but each dataset organizes the values in a different way:

```{r}
table1
```

```{r}
table2
```

```{r}
table3
```

```{r}
table4a
```

```{r}
table4b
```

**Three interrelated rules which make a daaset tidy:**
- Each variable must have its own column
- Each observation must have its own row
- Each value must have its own cell.

In this example, only table1 is tidy. It's the only representation where each column is a variable.

**Why ensure that your data is tidy? There are two main advantages:**

- There's a general advantage to picking one consistent way of storing data. If you have a consistent data structure, it's easier to learn the tools that work with it because they have an underlying uniformity.
- There's a specific advantage to placing variables in columns because it allows R's vectorized nature to shine. As you learned in "Useful Creation Functions" on Page 56 and "Useful summary functions" on page 66, most built-in R functions work with vector of values. That makes transforming tidy data feel particularly natural.

**Example showing how you might work with table1**

```{r}
# compute rate per 10,000
table1 %>%
  mutate(rate = cases / population * 10000)
```

```{r}
# compute cases per year
table1 %>%
  count(year, wt = cases)
```

```{r}
# visualize changes over time
ggplot(table1, aes(year, cases)) +
  geom_line(aes(group = country), color = "grey50") +
  geom_point(aes(color = country))
```

## Spreading and Gathering

**First Step** is always to figure out what the variables and observations are. 
**Second Step** is to resolve one of two common problems:
- One variable might be spread across multiple columns
- One observation might be scattered across multiple rows

### Gathering

A common problem in a dataset where some of the column names are not names of variables, but values of a variable.

```{r}
table4a
```

 - The column names 1999 and 2000 represent values of the year variable
 - each row represents two observations not one.

**Solution** is to gather those columns into a new pair of variables. For that, we need three parameters:
1. The *set of columns* that represent values, not variables. (1999 & 2000)
2. The name of the variable whose values from the column names. (called as "key",here it is "year")
3. The name of the variable whose values are spread over the cells. (called as "value", here it is "cases")


**Tidy table4a**

```{r}
table4a %>%
  gather(`1999`,`2000`, key = "year", value = "cases")
```

- nonsyntactic names should be surrounded in backticks

**Tidy table4b**

```{r}
table4b %>%
  gather(`1999`,`2000`, key = "year", value = "population")
```

**To combine tidied versions of table4a and table4b into a single tibble**

```{r}
tidy4a <- table4a %>%
  gather(`1999`,`2000`, key = "year", value = "cases")
tidy4b <- table4b %>%
  gather(`1999`,`2000`,key = "year", value = "population")

left_join(tidy4a,tidy4b)
```

### Spreading

Spreading is the opposite of gathering

```{r}
table2
```

**Identify two parameters before spreading**

1. The column that contains variable names, the key column.( type )
2. The column that contains values forms multiple variables, the value column ( count )

```{r}
spread(table2, key = type, value = count)
```

## Separating and Pull

### Separate

Separate pulls apart one column into multiple columns, by splitting wherever a separator character appears.

```{r}
table3
```

```{r}
table3 %>%
  separate(rate, into = c("cases", "population"))
```

```{r}
table3 %>%
  separate(rate, into = c("cases", "population"), sep = "/")
```

- default behavior of separate function -> leaves the type of the column as is.

```{r}
table3 %>%
  separate(
    rate,
    into = c("cases","population"),
    convert = TRUE
  )
```

```{r}
table3 %>%
  separate(year, into = c("century", "year"), sep = 2)
```

### Unite

unite() is the inverse of separate(): it combines multiple columns into a single column.

```{r}
table5
```

```{r}
table5 %>%
  unite(new, century, year)
```

```{r}
table5 %>%
  unite(new, century, year, sep = "")
```

## Missing Values

1. Explicitly i.e., flagged with NA
2. Implicitly i.e., Simply not present in the data

```{r}
stocks <- tibble(
  year = c(2015, 2015, 2015, 2015, 2016, 2016, 2016),
  qtr = c(1, 2, 3, 4, 2, 3, 4),
  return = c(1.88, 0.59, 0.35, NA, 0.92, 0.17, 2.66)
)
```

```{r}
stocks
```

Observations:

- return of the fourth quarter of 2015 is explicitly missing
- return of the first quarter of 2016 is implicitly missing




1. An explicit missing value is the presence of an absence
2. An implicit missing value is the absence of a presence

**We can make the implicit missing value explicit by putting years in the columns:**

```{r}
stocks %>%
  spread(key = year, value = return)
```

To convert explicit missing to implicit missing:

```{r}
stocks %>%
  spread(year, return) %>%
  gather(year, return, `2015`,`2016`, na.rm = TRUE)
```

```{r}
# To convert Implicit missing to Explicit missing
complete(stocks, year, qtr)
```

```{r}
treatment <- tribble(
  ~person, ~treatment, ~response,
  "Derrick Whitmore",1, 7,
  NA, 2, 10,
  NA, 3, 9,
  "Katherine Burke", 1, 4
)
```

```{r}
treatment
```

```{r}
treatment %>%
  fill(person)
```

## Case Study

```{r}
who
```

About:

The data comes from the 2014 World Health Organization Global Tuberculosis Report, available at (http://www.who.int/tb/country/data/download/en/)

```{r}
```


```{r}
colnames(who)
```
```{r}
head(who)
```
```{r}
summary(who)
```
```{r}
str(who)
```
```{r}
glimpse(who)
```
```{r}
who1<-who%>%gather(new_sp_m014:newrel_f65,key="key",value = "cases",na.rm=TRUE)
who1
```
```{r}
count(who1,key)
```
```{r}
who1<-who1%>%mutate(key=stringr::str_replace(key,"newrel","new_rel"))
who1
count(who1,key)
```

```{r}
colnames(who1)
(who1<-who1%>%select(-iso3,-iso2))
```

```{r}
(who1<-who1%>%separate(key,into=c("new","type","sexage")))
```
```{r}
who1<-who1%>%separate(sexage,into=c("sex","age"),sep=1)
```
```{r}
who1
```
```{r}
who %>%
 gather(code, value, new_sp_m014:newrel_f65, na.rm = TRUE) %>%
 mutate(
 code = stringr::str_replace(code, "newrel", "new_rel")
 ) %>%
 separate(code, c("new", "var", "sexage")) %>%
 select(-new, -iso2, -iso3) %>%
 separate(sexage, c("sex", "age"), sep = 1)

```

