---
title: "Data Transformation with dplyr Exercises"
author: "Sathish G"
date: "21/01/2020"
output: html_document
---

## Filter Rows filter()

### Load the libraries and DataSets:

```{r include=FALSE}
library(tidyverse)
library(nycflights13)
```

1. Find all flights that

a. Had an arrival delay of two or more hours

Since the arr_delay variable is measured in minutes, find flights with an arrival delay of 120 or more minutes.

        
```{r}
filter(flights, arr_delay >= 120)
```
        
b. Flew to Houston (IAH or HOU)

The flights that flew to Houston are those flights where the destination (dest) is either “IAH” or “HOU”.
        
```{r}
filter(flights, dest == "IAH" | dest == "HOU")
```
        
c. Were operated by United, American, or Delta

In the flights data set, the column carrier indicates the airline, but it uses two-character carrier codes. We can

find the carrier codes for the airlines in the airlines data set. Since the carrier code data set only has 16 rows,

and the names of the airlines in that data set are not exactly “United”, “American”, or “Delta”.

The carrier code for Delta is "DL", for American is "AA", and for United is "UA". Using these carriers codes, we

check whether carrier is one of those.
        
```{r}
airlines
filter(flights, carrier %in% c("AA", "DL", "UA"))
```
        
d. Departed in summer (July, August, and September)

The variable month has the month, and it is numeric. So, the summer flights are those that departed in months 7

(July), 8 (August), and 9 (September).
        
```{r}
filter(flights, month >= 7, month <= 9)
```
        
e. Arrived more than two hours late, but didn’t leave late

Flights that arrived more than two hours late, but didn’t leave late will have an arrival delay of more than 120

minutes (arr_delay > 120) and a non-positive departure delay (dep_delay <= 0).
        
```{r}
filter(flights, arr_delay > 120, dep_delay <= 0)
```
        
f. Were delayed by at least an hour, but made up over 30 minutes in flight

If a flight was delayed by at least an hour, then dep_delay >= 60. If the flight didn’t make up any time in the air,

then its arrival would be delayed by the same amount as its departure, meaning dep_delay == arr_delay, or

alternatively, dep_delay - arr_delay == 0. If it makes up over 30 minutes in the air, then the arrival delay must be

at least 30 minutes less than the departure delay, which is become dep_delay - arr_delay > 30.
        
```{r}
filter(flights, dep_delay >= 60, dep_delay - arr_delay > 30)
```
        
g. Departed between midnight and 6am (inclusive)

In dep_time, midnight is represented by 2400, not 0. We can see this from the below output
        
```{r}
summary(flights$dep_time)
```
```{r}
filter(flights, dep_time <= 600 | dep_time == 2400)

```
        
        
2. Another useful dplyr filtering helper is between(). What does it do? Can you use it to simplify the code needed to answer the previous challenges?

The expression between(x, left, right) is equivalent to x >= left & x <= right.

```{r}
filter(flights, between(month, 7, 9))
```


3. How many flights have a missing dep_time? What other variables are missing? What might these rows represent?

Find the rows of flights with a missing departure time (dep_time) using the is.na() function.

```{r}
filter(flights, is.na(dep_time))
```
The arrival time (arr_time) is also missing for these rows. These seem to be cancelled flights.

The output of the function summary() includes the number of missing values for all non-character variables.

```{r}
summary(flights)
```


4. Why is NA ^ 0 not missing? Why is NA | TRUE not missing? Why is FALSE & NA not missing? Can you figure out the general rule? (NA * 0 is a tricky counterexample!)

NA ^ 0 == 1 since for all numeric values x^0 =1

NA | TRUE is TRUE because anything or TRUE is TRUE. If the missing value were TRUE, then TRUE | TRUE == TRUE, and if

the missing value was FALSE, then FALSE | TRUE == TRUE.

The value of NA & FALSE is FALSE because anything and FALSE is always FALSE. If the missing value were TRUE, then

TRUE & FALSE == FALSE, and if the missing value was FALSE, then FALSE & FALSE == FALSE.

For NA & TRUE, the value is unknown since FALSE & TRUE== FALSE, but TRUE & TRUE == TRUE

```{r}
NA ^ 0
NA | TRUE
NA & FALSE
NA | FALSE
NA & TRUE
```


## Arrange Rows with arrange()

1. How could you use arrange() to sort all missing values to the start? (Hint: use is.na()).

The arrange() function puts NA values last.

```{r}
arrange(flights, dep_time) %>% tail()
```
```{r}
arrange(flights, desc(dep_time))
```
To put NA values first, we can add an indicator of whether the column has a missing value.

```{r}
arrange(flights, desc(is.na(dep_time)), dep_time)
```

2. Sort flights to find the most delayed flights. Find the flights that left earliest.

Find the most delayed flights by sorting the table by departure delay, dep_delay, in descending order.

```{r}
arrange(flights, desc(dep_delay))
```
The most delayed flight was HA 51, JFK to HNL, which was scheduled to leave on January 09, 2013 09:00. Note that the

departure time is given as 641, which seems to be less than the scheduled departure time. But the departure was

delayed 1,301 minutes, which is 21 hours, 41 minutes. 


3. Sort flights to find the fastest flights.

It mean that  “the flight with the shortest flight time”. We can use arrange to sort our data by the air_time

variable

```{r}
head(arrange(flights, air_time))
```
Another meaning for the “fastest flight” is the flight with the highest average ground speed. The ground speed is

not included in the data, so we have to calculate from the distance and air_time of the flight
```{r}
head(arrange(flights, desc(distance / air_time)))
```

4. Which flights travelled the longest? Which travelled the shortest?

To find the longest flight, sort the flights by the distance column in descending order.

```{r}
arrange(flights, desc(distance))
```
The longest flight is HA 51, JFK to HNL, which is 4,983 miles.

To find the shortest flight, sort the flights by the distance in ascending order, which is the default sort order.

```{r}
arrange(flights, distance)
```
The shortest flight is US 1632, EWR to LGA, which is only 17 miles. 

## Select columns with select()

1. Brainstorm as many ways as possible to select dep_time, dep_delay, arr_time, and arr_delay from flights.

Specify columns names as unquoted variable names.

```{r}
select(flights, dep_time, dep_delay, arr_time, arr_delay)
```

Specify column names as strings.
```{r}
select(flights, "dep_time", "dep_delay", "arr_time", "arr_delay")
```
Specify the column numbers of the variables.
```{r}
select(flights, 4, 6, 7, 9)
```

2. What happens if you include the name of a variable multiple times in a select() call?

The select() call ignores the duplication.

```{r}
select(flights, year, month, day, year, year)
```

3. What does the one_of() function do? Why might it be helpful in conjunction with this vector?

The one_of() function selects variables with a character vector rather than unquoted variable name arguments. 



```{r}
vars <- c("year", "month", "day", "dep_delay", "arr_delay")
select(flights, one_of(vars))
```

4. Does the result of running the following code surprise you? How do the select helpers deal with case by default? How can you change that default?

```{r}
select(flights, contains("TIME"))
```
The default behavior for contains() is to ignore case. 

To change the behavior add the argument ignore.case = FALSE.
```{r}
select(flights, contains("TIME", ignore.case = FALSE))
```


## Add New Variables with mutate() Exercises

1. Currently dep_time and sched_dep_time are convenient to look at, but hard to compute with because they’re not really continuous numbers. Convert them to a more convenient representation of number of minutes since midnight.

To get the departure times in the number of minutes, divide dep_time by 100 to get the hours since midnight and multiply by 60 and add the remainder of dep_time divided by 100. For example, 1504 represents 15:04

1504 %/% 100

Now, combine the hours (multiplied by 60 to convert them to minutes) and minutes to get the number of minutes after

midnight.

1504 %/% 100 * 60 + 1504 %% 100

 Midnight is represented by 2400, which would correspond to 1440 minutes. 
 
 After converting all the times to minutes after midnight, x %% 1440 will convert 1440 to zero while keeping all the
 
 other times the same
 

```{r}
flights_times <- mutate(flights,
  dep_time_mins = (dep_time %/% 100 * 60 + dep_time %% 100) %% 1440,
  sched_dep_time_mins = (sched_dep_time %/% 100 * 60 +
    sched_dep_time %% 100) %% 1440
)
# view only relevant columns
select(
  flights_times, dep_time, dep_time_mins, sched_dep_time,
  sched_dep_time_mins
)
```


2. Compare air_time with arr_time - dep_time. What do you expect to see? What do you see? What do you need to do to fix it?

air_time is the difference between the arrival (arr_time) and departure times (dep_time). In other words, air_time = arr_time - dep_time

To check this relationship, first need to convert the times to a previous exercise form

```{r}
flights_airtime <-
  mutate(flights,
    dep_time = (dep_time %/% 100 * 60 + dep_time %% 100) %% 1440,
    arr_time = (arr_time %/% 100 * 60 + arr_time %% 100) %% 1440,
    air_time_diff = air_time - arr_time + dep_time
  )
```


If air_time = arr_time - dep_time, there should be no flights with non-zero values of air_time_diff.

```{r}
nrow(filter(flights_airtime, air_time_diff != 0))
```
It meant that there are many flights for which air_time != arr_time - dep_time. 
May be  two reasons why air_time would not equal arr_time - dep_time

1. The flight passes midnight, so arr_time < dep_time. In these cases, the difference in airtime should be by 24 hours (1,440 minutes).

2. The flight crosses time zones, and the total air time will be off by hours (multiples of 60). All flights in flights departed from New York City and are domestic flights in the US

Given the time-zones in the US, the differences due to time-zone should be 60 minutes (Central) 120 minutes (Mountain), 180 minutes (Pacific), 240 minutes (Alaska), or 300 minutes (Hawaii).

3. Compare dep_time, sched_dep_time, and dep_delay. How would you expect those three numbers to be related?


The departure delay (dep_delay) to be equal to the difference between scheduled departure time (sched_dep_time), and actual departure time (dep_time), dep_time - sched_dep_time = dep_delay.

As with the previous question, the first step is to convert all times to the number of minutes since midnight. The column, dep_delay_diff, is the difference between the column, dep_delay, and departure delay calculated directly from the scheduled and actual departure times.
```{r}
flights_deptime <-
  mutate(flights,
    dep_time_min = (dep_time %/% 100 * 60 + dep_time %% 100) %% 1440,
    sched_dep_time_min = (sched_dep_time %/% 100 * 60 +
      sched_dep_time %% 100) %% 1440,
    dep_delay_diff = dep_delay - dep_time_min + sched_dep_time_min
  )
```


Is dep_delay_diff equal zero for all rows

```{r}
nrow(filter(flights_deptime, dep_delay_diff != 0))
```

No. Unlike the last question, time zones are not an issue since because we only considering departure times.

The discrepancies may be because a flight was scheduled to depart before midnight, but was delayed after midnight.

4. Find the 10 most delayed flights using a ranking function. How do you want to handle ties? Carefully read the documentation for min_rank().

```{r}
flights_delayed <- mutate(flights, 
                          dep_delay_min_rank = min_rank(desc(dep_delay)),
                          dep_delay_row_number = row_number(desc(dep_delay)),
                         )
flights_delayed <- filter(flights_delayed, dep_delay_min_rank <= 10 )
flights_delayed <- arrange(flights_delayed, dep_delay_min_rank)
flights_delayed
```


5. What does 1:3 + 1:10 return? Why?

```{r}
1:3 + 1:10
```


When adding two vectors, R recycles the shorter vector’s values to create a vector of the same length as the longer vector. The code also raises a warning that the shorter vector is not a multiple of the longer vector

6. What trigonometric functions does R provide?

```{r}

```

## Grouped Summaries with summarize()

```{r}
not_cancelled <- flights %>%
                    filter(!is.na(dep_delay), !is.na(arr_delay))
```

1. Brainstorm at least 5 different ways to assess the typical delay characteristics of a group of flights. Consider the following scenarios:
        ◦ A flight is 15 minutes early 50% of the time, and 15 minutes late 50% of the time.
        ◦ A flight is always 10 minutes late.
        ◦ A flight is 30 minutes early 50% of the time, and 30 minutes late 50% of the time.
        ◦ 99% of the time a flight is on time. 1% of the time it’s 2 hours late.

Which is more important: arrival delay or departure delay?

In many scenarios, arrival delay is more important. In most cases, being arriving late is more costly to the passenger since it could affect the next stages of their travel, such as connecting flights or scheduled meetings.

If a departure is delayed without affecting the arrival time, this delay will not affects plans or the total time spent traveling. 


3. Come up with another approach that will give you the same output as not_cancelled %>% count(dest) and not_cancelled %>% count(tailnum, wt = distance) (without using count()).

```{r}
not_cancelled <- flights %>%
  filter(!is.na(dep_delay), !is.na(arr_delay))
```
```{r}
not_cancelled %>% 
  count(dest)
```
Instead of using the count() function, we can combine the group_by() and summarise() 
```{r}
not_cancelled %>%
  group_by(dest) %>%
  summarise(n = length(dest))
```
    
4. Our definition of cancelled flights (is.na(dep_delay) | is.na(arr_delay) ) is slightly suboptimal. Why? Which is the most important column?

If a flight never departs, then it won’t arrive. A flight could also depart and not arrive if it crashes, or if it is redirected and lands in an airport other than its intended destination. So the most important column is arr_delay, which indicates the amount of delay in arrival.

```{r}
filter(flights, !is.na(dep_delay), is.na(arr_delay)) %>%
  select(dep_time, arr_time, sched_arr_time, dep_delay, arr_delay)
```


5. Look at the number of cancelled flights per day. Is there a pattern? Is the proportion of cancelled flights related to the average delay?


```{r}
cancelled_per_day <- 
  flights %>%
  mutate(cancelled = (is.na(arr_delay) | is.na(dep_delay))) %>%
  group_by(year, month, day) %>%
  summarise(
    cancelled_num = sum(cancelled),
    flights_num = n(),
  )
```

The relationship between the number of cancelled flights per day and the total number of flights in a day. There should be an increasing relationship for two reasons. First, if all flights are equally likely to be cancelled, then days with more flights should have a higher number of cancellations. Second, it is likely that days with more flights would have a higher probability of cancellations because congestion itself can cause delays and any delay would affect more flights, and large delays can lead to cancellations.
```{r}
ggplot(cancelled_per_day) +
  geom_point(aes(x = flights_num, y = cancelled_num)) 
```
Plotting flights_num against cancelled_num shows that the number of flights cancelled increases with the total number of flights.

6. Which carrier has the worst delays? Challenge: can you disentangle the effects of bad airports vs. bad carriers? Why/why not? (Hint: think about flights %>% group_by(carrier, dest) %>% summarise(n()))

```{r}
flights %>%
  group_by(carrier) %>%
  summarise(arr_delay = mean(arr_delay, na.rm = TRUE)) %>%
  arrange(desc(arr_delay))
```

What airline corresponds to the "F9" carrier code?
```{r}
filter(airlines, carrier == "F9")
```

7. What does the sort argument to count() do. When might you use it?

```{r}
flights %>%
  count(dest, sort = TRUE)
```


## Grouped Mutates and Filters

1. Refer back to the lists of useful mutate and filtering functions. Describe how each operation changes when you combine it with grouping.

```{r}

```


2. Which plane (tailnum) has the worst on-time record?

```{r}

```


3. What time of day should you fly if you want to avoid delays as much as possible?

```{r}

```


4. For each destination, compute the total minutes of delay. For each flight, compute the proportion of the total delay for its destination.

```{r}

```


5. Delays are typically temporally correlated: even once the problem that caused the initial delay has been resolved, later flights are delayed to allow earlier flights to leave. Using lag(), explore how the delay of a flight is related to the delay of the immediately preceding flight.

```{r}

```


6. Look at each destination. Can you find flights that are suspiciously fast? (i.e. flights that represent a potential data entry error). Compute the air time a flight relative to the shortest flight to that destination. Which flights were most delayed in the air?

```{r}

```


7. Find all destinations that are flown by at least two carriers. Use that information to rank the carriers.

```{r}

```


8. For each plane, count the number of flights before the first delay of greater than 1 hour.

```{r}

```