---
title: "r4ds-ex11-Grouped Summaries with summarize() Exercises"
author: "Sathish G"
date: "01/10/2019"
output: html_document
---

## load libraries and datasets

```{r include=FALSE}
library(tidyverse)
library(nycflights13)
```


## summarise() Exercises

1. Brainstorm at least 5 different ways to assess the typical delay characteristics of a group of flights. Consider the following scenarios:
    a. A flight is 15 minutes early 50% of the time, and 15 minutes late 50% of the time
    
```{r}
not_cancelled <- flights %>% filter(!is.na(dep_delay), !is.na(arr_delay))

not_cancelled %>% group_by(tailnum) %>% summarise(delay_perc = mean(arr_delay == 15)) %>% filter(delay_perc == 0.50)

not_cancelled %>% group_by(tailnum) %>% summarise(delay_perc = mean(dep_delay == -15)) %>% filter(delay_perc == 0.50)

```
    
    
    b. A flight is always 10 minutes late
    
```{r}
not_cancelled %>% group_by(tailnum) %>% summarise(delay_perc = mean(dep_delay == -10)) %>% filter(delay_perc == 1.0)
```
    
    
    c. A flight is 30 minutes early 50% of the time, and 30 minutes late 50% of the time.
    
```{r}
not_cancelled %>% group_by(tailnum) %>% summarise(delay_perc = mean(dep_delay == 30)) %>% filter(delay_perc == 0.50)

not_cancelled %>% group_by(tailnum) %>% summarise(delay_perc = mean(arr_delay == 30)) %>% filter(delay_perc == 0.50)
```
    
    
    d. 99% of the time a flight is on time. 1% of the time it's 2 hours late.
    
```{r}
not_cancelled %>% group_by(tailnum) %>% summarise(delay_perc = mean(dep_delay == 120)) %>% filter(delay_perc == 0.01)
```
    
    
    Which is more important: arrival delay or departure delay ?
    
    both arrival delay as well as departure delay is important depends on the situation or the question been asked.
    
2. Come up with another approach that will give you the same output as not_cancelled %>% count(dest) and not_cancelled %>% count(tailnum,wt = distance) (without using count())

```{r}
not_cancelled %>% count(dest)
```


```{r}
not_cancelled %>% group_by(dest) %>% summarise(n = n())
```

The above results can be compared to check the similarity / differences.

```{r}
not_cancelled %>% count(tailnum,wt = distance)
```


```{r}
not_cancelled %>% group_by(tailnum) %>% summarise(n = sum(distance))
```


3. Our definition of cancelled flights (is.na(dep_delay) | is.na(arr_delay)) is slightly suboptimal. Why? Which is the most important column?

The most important column would be arr_delay.

```{r}
flights %>% filter(is.na(dep_delay),!is.na(arr_delay)) %>% select(dep_time,arr_time,sched_arr_time,dep_delay,arr_delay)
```


4. Look at the number of cancelled flights per day. Is there a pattern ? Is the proportion of cancelled flights related to the average delay ?

```{r}
flights %>% mutate(cancelled = (is.na(arr_delay) | is.na(dep_delay))) %>% group_by(year,month,day) %>% summarise(cancelled_mean = mean(cancelled), dep_delay_mean = mean(dep_delay, na.rm = TRUE)) %>% ggplot(aes(x = dep_delay_mean, y = cancelled_mean)) + geom_point() + geom_smooth()
```


5. Which carrier has the worst delays ? Challenge: can you disentangle the effects of bad airports vs bad carriers? Why/Why not ? (Hint: think about flights %>% group_by(carrier,dest) %>% summarise(n()))

which carrier has the worst delays ?

```{r}
flights %>% group_by(carrier) %>% summarise(delay = mean(arr_delay,na.rm = TRUE)) %>% arrange(desc(delay))
```

The first carrier has the most average delay.

```{r}
airlines %>% filter(carrier == "F9")
```


6. What does the sort argument to count() do. When might you use it?

```{r}
flights %>% group_by(carrier) %>% count(dep_delay, sort = TRUE)
```
