---
title: "r4ds-ex7"
author: "Sathish G"
date: "30/08/2019"
output: html_document
---

# Loading Data

Before starting the exercises load the packages and data sets that are needed to complete the exercises. Here, you can either load `dplyr` package directly or `tidyverse`

```{r}
library(ggplot2)
library(dplyr)
```

### Loading Data Sets

```{r}
library(nycflights13)
```

# filter() Exercises

1. Find all flights that:
    a. Had an arrival delay of two or more hours
    
```{r}
names(flights)
```
    
```{r}
filter(flights,arr_delay>=120)
```
    
    
    b. Flew to Houston (IAH or HOU)
    
```{r}
names(flights)
```
    
```{r}
flights %>% select(dest) %>% group_by(dest)
```
    
    
```{r}
filter(flights, dest=="IAH"| dest=="HOU")
```
    
    c. Were operated by United, American, or Delta
    
```{r}
names(flights)
```
    
```{r}
head(flights$carrier)
```
    
```{r}
filter(flights,carrier == "UA" | carrier == "AA" | carrier == "DL")
```
    
    
    d. Departed in summer (July, August, and September)
    
```{r}
names(flights)
```
    
```{r}
head(flights$month)
```
    
```{r}
filter(flights,month %in% c(7,8,9))
```
    
    
    e. Arrived more than two hours late, but didn't leave late
    
```{r}
names(flights)
```
    
```{r}
filter(flights,arr_delay>120, dep_delay<=0)
```
    
    f. Were delayed by at least an hour, but made up over 30 minutes in flight
    
```{r}
filter(flights,dep_delay>=60 & dep_delay - arr_delay>30)
```
    
    g. Departed between midnight and 6 a.m. (inclusive)
    
```{r}
names(flights)
```
    
```{r}
filter(flights,dep_time==2400 | dep_time<=600)
```


2. Another useful dplyr filtering helper is between(). What does it do? Can you use it to simplify the code needed to answer the previous challenges?

```{r}
filter(flights,between(month,7,9))
```


```{r}
filter(flights,between(dep_time,0,360))
```


3. How many flights have a missing dep_time? What other variables are missing? What might these rows represent ?

Answering 3.1 Question `How many flights have a missing dep_time ?`

```{r}
filter(flights,is.na(dep_time))
```

```{r}
filter(flights,is.na(dep_time)) %>% count()
```

Answering 3.2 Question `What other variables are missing ?`

```{r}
filter_all(flights,any_vars(is.na(.)))
```

`dep_delay and arr_time are the other variables which has missing values (NA's)`

4. Why is NA ^ 0 not missing ? Why is NA | TRUE not missing ? Why is FALSE & NA not missing? Can you figure out the general rule? (NA * 0 is a tricky counterexample!)

4.1 Why is NA ^ 0 not missing ?

because, mathematically speaking, anything power 0 gives 1.

```{r}
NA ^ 0
```

4.2 Why is NA | TRUE not missing ?

because, | requires any one condition to be TRUE to yeild the result TRUE.

```{r}
NA | TRUE
```

whereas NA | NA will yield the result NA, as we don't know either one of the condition to be TRUE.

```{r}
NA | NA
```

4.3 Why is FALSE & NA not missing ?

as & results in FALSE when one condition is FALSE

```{r}
NA & FALSE
```

4.4 Can you figure out the general rule ?

```{r}
NA * 0
```

NA represents unknown value. As the value of NA is unknown, with one unknown value in some expressions, it is harder to predict the results for those expressions.
