---
title: "Add New Variables with mutate() Exercises"
author: "Sathish G"
date: "30/09/2019"
output: html_document
---

## load libraries and datasets

```{r include=FALSE}
library(tidyverse)
library(nycflights13)
```


## mutate() Exercises

1. Currently dep_time and sched_dep_time are convenient to look at, but hard to compute with because they're not really continuous numbers. Convert them to a more convenient representation of number of minutes since midnight.

Set a duplicate data set to work with

```{r}
flights_sml <- select(flights,year:day,dep_time,sched_dep_time)
flights_sml
```

Create time variables denoting number of minutes since midnight for variables dep_time and sched_dep_time

```{r}
flights_sml <- mutate(flights_sml,
                      dep_time_minutes = (dep_time %/% 100 * 60 + dep_time %% 100) %% 1440,
                      sched_dep_time_minutes = (sched_dep_time %/% 100 * 60 + sched_dep_time %% 100) %% 1440
                        )
flights_sml
```

2. Compare air_time with arr_time - dep_time. What do you expect to see ? What do you see? What do you need to do to fix it?

```{r}

timeToMinutes <- function(timeInhhmm){
  return (timeInhhmm %/% 100 * 60 + timeInhhmm %% 100) %% 1440
}

flights_sml <- select(flights,air_time,arr_time,dep_time)
flights_sml <- mutate(flights_sml,
                      arr_time_minutes = timeToMinutes(arr_time),
                      dep_time_minutes = timeToMinutes(dep_time),
                      diff = arr_time_minutes - dep_time_minutes)
flights_sml
```

I see that the air_time and (arr_time - dep_time) difference is different.

3. Compare dep_time, sched_dep_time, and dep_delay. How would you expect those three numbers to be related?

```{r}
flights_dep <- select(flights,dep_time,sched_dep_time,dep_delay)
flights_dep <- mutate(flights_dep,cal_dep_delay = timeToMinutes(dep_time) -timeToMinutes(sched_dep_time))
flights_dep
```

4. Find the 10 most delayed flights using a ranking function. How do you want to handle ties ? Carefully read the documentation for min_rank().

```{r}
arrange(flights,desc(dep_delay)) %>% mutate(rank = min_rank(desc(dep_delay))) %>% head(10) %>% select(rank,dep_delay)
```


5. What does 1:3 + 1:10 return ? why ?

```{r}
1:3 + 1:10
```

This is because, when doing arithmetic between two vectors, the vectors are supposed to be of same length. but it doesn't mean that we made any error.

6. What trignometric functions does R provide ?

```{r}
help("Trig")
```

cos(x)
sin(x)
tan(x)
acos(x)
asin(x)
atan(x)
atan2(y,x)
cospi(x)
sinpi(x)
tanpi(x)

Where x & y -> numeric or complex vectors.

Examples given in documentation:

```{r}
x <- seq(-3, 7, by = 1/8)
tx <- cbind(x, cos(pi*x), cospi(x), sin(pi*x), sinpi(x),
               tan(pi*x), tanpi(x), deparse.level=2)
op <- options(digits = 4, width = 90) # for nice formatting
head(tx)
tx[ (x %% 1) %in% c(0, 0.5) ,]
options(op)
```
