---
title: "r4ds-ex9-Select Columns with select() Exercises"
author: "Sathish G"
date: "30/09/2019"
output: html_document
---

## load the libraries and datasets

```{r include=FALSE}
library(tidyverse)
library(nycflights13)
```


## select() Exercises

1. Brainstorm as many ways as possible to select dep_time, dep_delay, arr_time, and arr_delay from flights.

```{r}
select(flights,dep_time,dep_delay,arr_time,arr_delay)
```

```{r}
select(flights,starts_with("dep"),starts_with("arr"))
```


2. What happens if you include the name of a variable multiple times in a select() call ?

```{r}
select(flights,arr_time,arr_time,dep_time,arr_time)
```


3. What does the one_of() function do ? Why might it be helpful in conjunction with this vector ?

vars <- c("year","month","day","dep_delay","arr_delay")

```{r}
vars <- c("year","month","day","dep_delay","arr_delay")
select(flights,one_of(vars))
```


4. Does the result of running the following code surprise you ? How do the select helpers deal with case by default ? How can you change that default ?

select (flights,contains("TIME"))

```{r}
select (flights,contains("TIME"))
```

```{r}
select(flights,contains("TIME",ignore.case = FALSE))
```

```{r}
select(flights,contains("TIME",ignore.case = TRUE))
```
