---
title: "r4ds-ex12-Grouped Mutates and Filters Solutions"
author: "Sathish G"
date: "01/10/2019"
output: html_document
---

## load libraries and datasets

```{r include=FALSE}
library(tidyverse)
library(nycflights13)
```


## Grouped Mutates(and Filters) Exercises

1. Refer back to the table of useful mutate and filtering functions. Describe how each operation changes when you combine it with grouping.

2. Which plane (tailnum) has the worst on-time record ?

```{r}
flights %>% group_by(tailnum) %>% summarise(mean_arr_delay = mean(arr_delay)) %>% filter(min_rank(desc(mean_arr_delay)) <= 1)
```


3. What time of day should you fly if you want to avoid delays as much as possible ?

```{r}
flights %>% filter(!(is.na(arr_delay))) %>% group_by(hour) %>% summarise(mean_arr_delay = mean(arr_delay)) %>% arrange(mean_arr_delay)
```


4. For each destination, compute the total minutes of delay. For each flight, compute the proportion of the total delay for its destination.

```{r}
flights %>% filter(!(is.na(arr_delay)) & arr_delay > 0) %>% group_by(dest) %>% mutate(sum_arr_delay = sum(arr_delay), proportion_arr_delay = arr_delay/sum_arr_delay) %>% select(dest,sum_arr_delay,proportion_arr_delay)
```


5. Delays are typically temporally correlated: even once the problem that caused the initial delay has been resolved, later flights are delayed to allow earlier flights to leave. Using lag() explores how the delay of a flight is related to the delay of the immediately preceding flight.

```{r}
lag_flights <- flights %>% arrange(year,month,day,dep_time) %>% mutate(lag_dep_delay = lag(dep_delay))

lag_flights <- lag_flights %>% filter(!(is.na(dep_delay)), !(is.na(lag_dep_delay)))

lag_flights %>% group_by(origin) %>% summarise(mean_dep_delay = mean(dep_delay), mean_lag_dep_delay = mean(lag_dep_delay),diff = mean_dep_delay - mean_lag_dep_delay)

```


6. Look at each destination. Can you find flights that are suspiciously fast ? (That is, flights that represent a potential data entry error.) Compute the air time of a flight relative to the shortest flight to that destination. Which flights were most delayed in the air?

```{r}
fast_flights <- flights %>% mutate(speed = distance/air_time*60) %>% arrange(desc(speed))
fast_flights %>% group_by(tailnum,dest) %>% summarize(mean = mean(speed,na.rm = TRUE)) %>% arrange(desc(mean))
```

```{r}
flights %>% group_by(tailnum,dest) %>% summarise(mean_distance = mean(distance), mean_air_time = mean(air_time), mean_speed = mean(mean_distance/mean_air_time)) %>% arrange(mean_distance,mean_speed) %>% filter(!(is.na(tailnum)))
```


7. Find all destinations that are flown by at least two carriers. Use that information to rank the carriers.

```{r}
flights %>% group_by(dest) %>% summarize(carrier_count = n_distinct(carrier)) %>% filter(carrier_count > 1) %>% mutate(rank = min_rank(desc(carrier_count))) %>% arrange(rank)
```
