---
title: 'r4ds-ex8-Data Transformation with dplyr : Arrange Rows with arrange()'
author: "Sathish G"
date: "23/08/2019"
output: html_document
---
```{r}
# load the tidyverse library
library(tidyverse)
# load the nycflights13 dataset
library(nycflights13)
```


1. How could you use arrange() to sort all missing values to the start ? (Hint: use is.na())

```{r}
df <- tibble(x = c(5,2,NA,NA,3,8,7,NA,1,1,NA))
arrange(df,x)
```

```{r}
df <- tibble(x = c(5,2,NA,NA,3,8,7,NA,1,1,NA))
arrange(df,desc(is.na(x)))
```

2. Sort flights to find the most delayed flights. Find the flights that left earliest.

2.1 Sort flights to find the most delayed flights.

```{r}
arrange(flights,desc(dep_delay))
```

2.2 Find the flights that left earliest.

```{r}
filter(flights,dep_delay<0) %>% arrange()
```


3. Sort flights to find the fastest flights.

```{r}
arrange(flights,air_time)
```

`because less air_time of a flight indicates that the flight travels less time but it doesn't tell that it is faster than other flights.so, to get speed, let us divide distance by air_time`

```{r}
arrange(flights,desc(distance/air_time*60))
```


4. Which flights traveled the longest ? Which traveled the shortest ?

to find the flights that traveled longest distance

```{r}
arrange(flights,desc(distance));
```

to find the flights that traveled shortest distance

```{r}
arrange(flights,distance)
```
