---
title: "dplyr"
author: "Sathish G"
date: "08/08/2019"
output: html_document
---

## Workflow: Basics

### 1. Coding Basics

```{r}
  1 / 200 * 30
```

```{r}
(59 + 73 + 2) / 3
```

```{r}
sin(pi/2)
```

```{r}
x <- 3 * 4
```

### 2. What's in a name

i_use_snake_case
otherPeopleUseCamelCase
some.people.use.periods
And_aFew.People_RENOUNCEconvention

```{r}
this_is_a_really_long_name <- 2.5
```

> ^ - exponent operator

```{r}
r_rocks <- 2 ^ 3
```

> Variables are case sensitive
> mistyped variable names results in an error

```{r eval=FALSE, include=FALSE}
r_rock
```

> variables should use the same case

```{r eval=FALSE, include=FALSE}
R_rocks
```

### 3. Calling Functions

```{r}
# function_name(arg1,arg2,arg3,...argn)
seq(1,10)
seq(from = 3,to = 15,by = 2)
```
n
```{r}
x <- "hello world"
typeof(x)  # returns the type of object
```

```{r}
(y <- seq(1,10, length.out = 7))
```

```{r}
(y <- seq(1,10,length.out = 9))
```

## Data Transformation with dplyr

### 1. Introduction

> dplyr : package of tidyverse used for manipulating the data or transforming the data


### 2. Prerequisites

```{r include=FALSE}
library(tidyverse)
library(nycflights13)
```

### 3. nycflights13

> nycflights13::flights dataframe contains all 336,776 flights that departed from New York City in 2013.

> Data comes from The US Bureau of Transportation Statistics, and is documented in ?flights


```{r}
nycflights13::flights
```

```{r}
flights
```

> dimensions of flights

```{r}
dim(flights)
```

> head of dataframe flights

```{r}
head(flights, n =6)
```

```{r}
colnames(flights)
```

Type of each variable :
  - int stands for integers
  - dbl stands for doubles, or real numbers
  - chr stands for character vectors or strings
  - dttm stands for date-times (a date + a time)


### 4. dplyr basics

dplyr functions:

  - filter observations - filter()
  - arrange observations - arrange()
  - select variables - select()
  - create new variables of existing variables - mutate(), transmute()
  - summarize variables and observations - groupby(), summarize()

All verbs work similarly:

  - The first argument is a data frame
  - The subsequent arguments describe what to do with the data frame, using the variable names (without quotes)
  - The result is a new data frame.

### 5. Filter Rows with filter()

  > filter() allows you to subset observations based on their values.

  > Question: Filter all the flights which departed in 1st of January.?

  > To list the count of each category in a categorical variable

```{r}
count(flights, month)
```


```{r}
filter(flights, month == 1, day == 1)
```

  > If you want to save the result
  
```{r}
jan1 <- filter(flights, month == 1, day == 1)
```
  
  > R either prints out the results or saves them to a variable. If we want to do both, wrap the assignment in parentheses:
  
```{r}
(dec25 <- filter(flights, month == 12, day == 25))
```
  
## Date : 16.08.2019
#### 5.1 Comparisons

  > Using comparison operators with filter()
  > comparison operators : 
      - greater than (>)
      - greater than or equal to (>=)
      - less than (<)
      - less than or equal to (<=)
      - not equal (!=)
      - equal (==)
      
  > Using = instead of == is a mistake

```{r}
filter(flights, month != 1) %>% group_by(month) %>% summarize(count = n())
```


```{r eval=FALSE, include=FALSE}
filter(flights, month = 1)
```

  > when using == with floating-point numbers, remember that the computer use finite precision arithmetic
  
```{r}
sqrt(2) ^ 2 == 2
```
  
```{r}
1 / 49 * 49 == 1
```
  
  
  > every number we see is an approximation.
  > Instead of ==, we can use near() function in those cases
  
```{r}
near(sqrt(2) ^ 2,2)
```
  

```{r}
near(1 / 49 * 49,1)
```

#### 5.2 Logical Operators

  > Multiple arguments to filter() are combined with "and:
        - every expression must be true in order for a row to be included in the output.
  > Other types of combinations : 
        - & is "and"
        - | is "or"
        - ! is "not"
        
  > Question: Find all flights that departed in November or December ?

```{r}
filter(flights,!(month == 10))
```


```{r}
filter(flights, month == 11 | month == 12)
```
  
  > The following expression is wrong !!
  
```{r}
filter(flights, month == 11 | 12)
```
  
  > Shorthand writing in filter()
  0
```{r}
nov_dec <- filter(flights, month %in% c(11,12))
```
  
  > De Morgan's Law
      - !(x & y) is the same as !x | !y
      - !(x | y) is the same as !x & !y
  
  > For Example:
  > Question: Find flights that weren't delayed (on arrival or departure) by more than two hours
  
```{r}
filter(flights, !(arr_delay > 120 | dep_delay > 120))
```
  
  > or
  
```{r}
filter(flights, arr_delay <= 120, dep_delay <= 120)
```
  
## Date : 19.08.2019
#### 5.3 Missing Values
  
  > NA - "Not Availables"
  > NA represents an unknown value, so missing values are "contagious"
  > Almost any operation in
  
  
```{r}
NA > 5
```
  
```{r}
10 == NA
```
  
```{r}
NA + 10
```

```{r}
NA / 2
```

```{r}
NA == NA
```

  > An example to understand why this is true ?
  
  > Let x be Mary's age. We don't know how old she is.
  
```{r}
x <- NA
```
  
  > Let y be John's age. We don't know how old she is.
  
```{r}
y <- NA
```
  
  > Are John and Mary the same age ?
  
```{r}
x == y
```
  
  > Answer : We don't know!
  
  > If you want to determine if a value is missing ?
  
```{r}
is.na(x)
```
  
  > by default, filter() excludes NA
  > we can include NA by explicit mention
  
```{r}
df <- tibble(x = c(1,NA,3))
df
```
  
```{r}
filter(df, x > 1)
```
  
```{r}
filter(df, is.na(x) | x > 1)
```
  

### 6. Arrange Rows with arrange()

  > arrange() works similarly to filter() except that instead of selecting rows, it changes their order.
  > It takes a data frame and a set of column names to order by
  > If you provide more than one column name, each individual column will be used to break ties in the values of preceding columns.
  
  > Question: Arrange the flights with respect to year, month , day ?
  
  
```{r}
# arrange flights based on month in ascending order
arrange(flights,month)
```
  
  
```{r}
# arrange flights based on month in descending order
arrange(flights,desc(month))
```

  
```{r}
arrange(flights, year, month, day)
```
  
```{r}
arrange(flights, desc(dep_time),desc(sched_dep_time))
# please scroll through  521 - 530
```
  
  
  > Use desc() to reorder by a column in descending order
  > Question: Arrange flights based on the arr_delay from greater delays to smaller delays.
  
```{r}
arrange(flights, desc(arr_delay))
```
  
  > Missing values are always sorted at the end
  
```{r}
df <- tibble(x = c(5, 2, NA))
arrange(df,x)
```
  
```{r}
arrange(df, desc(x))
```


### 7. Select columns with select()

  > Challenge: datasets with hundreds or even thousands of variables
  > We can only select variables in which we are interested in.
  
  > Question: Select variables year, month, day from the flights data set ?
  
```{r}
select(flights, year, month, day)
```
  
  > Question: Select all columns between year and day (inclusive) ?
  
```{r}
flights
```
  
  
```{r}
select(flights, year:dep_time)
```
  
  > Question: Select all columns between year and dep_time(inclusive) except day
  
```{r}
select(flights, year:dep_time, -day)
```
  
  
  > Question: Select all columns except those from year to day (inclusive)
  
```{r}
select(flights, -(year:day))
```
  
  > Helper functions within select():
      - starts_with("abc") matches names that begin with "abc"
      
```{r}
select(flights, starts_with("dep"))
```
      
      - ends_with("xyz") matches names that end with "xyz"
      
```{r}
select(flights, ends_with("time"))
```
      
      - contains("ijk") matches names that contain "ijk"
      
```{r}
select(flights, contains("dep"))
```
      
      - matches("(.)\\1") selects variables that match a regular expression. This one matches any variables that contain repeated characters. 
      
```{r}
select(flights, matches(".*arr_time"))
```
      
```{r}
select(flights, matches(".+arr_time"))
```
      
      
      - num_range("x", 1:3) matches x1, x2, x3
      
  > select() can be used to rename variables, but it's rarely useful because it drops all of the variables not explicitly mentioned.
  > Instead use rename(), which is a variant of select() that keeps all the variables that aren't explicitly mentioned.
  
```{r}
names(flights)
rename(flights, tail_num = tailnum)
```
  
```{r}
select(flights, tailnum,origin, dest, everything())
```
    
```{r}
flights
```
    
```{r}
# discussion on pipeline symbol
filter(flights, arr_delay >= 120) %>% select(tailnum,day,arr_delay) %>% arrange(desc(day), desc(arr_delay))
```
    
    
### 8. Add New Variables with mutate()

  > mutate() adds new columns at the end of your dataset.
  
```{r}
(flights_sml <- select(flights, year:day,ends_with('delay'),distance, air_time))
```
  
  > adds new column called gain which has the value of arr_delay - dep_delay
  > gain value signifies the delay during the flight
  > negative value -> speed (fast) of the travel
  > positive value -> speed (slow)
  > 0 -> actual speed or expected speed of the flight
  
```{r}
mutate(flights_sml, gain = arr_delay - dep_delay)
```
  

```{r}
mutate(flights_sml, gain = arr_delay - dep_delay, speed = distance/air_time * 60) %>% arrange(desc(speed))
```

```{r}
transmute(flights,
          gain = arr_delay - dep_delay,
          hours = air_time / 60,
          gain_per_hour = gain / hours)
```

#### Useful Creation functions

1. Arithmetic Operators 
    - +
    - -
    - *
    - /
    
```{r}
12 / 5
```
    
    
    - ^
    
2. Modular Arithmetic
    - %/% (Integer Division)
    
```{r}
12 %/% 5
```
    
    - %% (Remainder)
    
```{r}
1 %% 3
```
    
```{r}
transmute(flights, 
          dep_time,
          hour = dep_time %/% 100,
          minute = dep_time %% 100)
```
    
3. Logs log(), log2(), Log10()
    > Logarithms are an incredibly useful tranformation for dealing with data that ranges across multiple orders of magnitude.
    
4. Offsets

  > lag(x) and lead(x) allow you to refer to leading or lagging values.
  > This allows you to compute running differences (eg. x - lag(x))

```{r}
(x <- 1:10)
lag(x)
lead(x)
# lag and lead are useful in calculating the running differences
x - lag(x)
```

```{r}
(y <- c(4,10,12))
lag(y)
lead(y)
y - lag(y)
lead(y) - y
```

5. Cumulative and rolling aggregates

```{r}
(x <- 1: 10)
cumsum(x)
```

```{r}
cummean(x)
# first value - same value
# second value - sum of first two values / 2
# third value - sum of first three values / 3
```

```{r}
# minimum value of first r values
# where r - 1 value, 2 values, 3 values, etc.
cummin(x)
cummax(x)
y
cummin(y)
cummax(y)
```

6. Logical Comparisons

    - <
    - <= 
    - >
    - >= 
    - !=
    
7. Ranking

```{r}
y <- c(1,2,2,NA, 3, 4)
y
# min_rank -> give equal rank for equal values
# 
min_rank(y)
min_rank(desc(y))
row_number(y)
dense_rank(y)
percent_rank(y)
cume_dist(y)
```

### 9. Grouped Summaries with summarize()

  > summarise() is not terribly useful unless we pair it with group_by()
  > summarise collapses a data frame to a single row
  > all the summary function have na.rm argument
  
```{r}
select(flights, dep_delay, everything())
summarise(flights, sum = sum(dep_delay,na.rm = TRUE),delay = mean(dep_delay, na.rm = TRUE), avg = sum/328521)
```
  

#### 9.1 Combining Multiple Operations with the Pipe

```{r}
select(flights, tailnum,origin,dest)
```

```{r}
count(flights,dest)
group_by(flights,dest) %>% summarise(mean = mean(arr_delay, na.rm = TRUE))
```

```{r}
flights %>%
  filter(dest == 'HNL') %>%
  summarise(count = n())
```

```{r}
count(flights,dest) %>% filter(n<=20)
```



```{r}
(by_dest <- group_by(flights, dest))
(delay <- summarise(by_dest,
                   count = n(),
                   dist = mean(distance, na.rm = TRUE),
                   delay = mean(arr_delay, na.rm = TRUE)
                   ))
(delay <- filter(delay, count>20, dest!="HNL"))

ggplot(data = delay, mapping = aes(x = dist, y = delay)) +
  geom_point(aes(size = count), alpha = 1/3) +
  geom_smooth(se = FALSE)

```


1. Group flights by destination
2. Summarise to compute distance, average delay, and number of flights
3. Filter too remove noisy points and Honolulu airport, which is almost twice as far away as the next closest airport.

```{r}
delays <- flights %>%
  group_by(dest) %>%
  summarize(
    count = n(),
    dist = mean(distance, na.rm = TRUE),
    delay = mean(arr_delay, na.rm = TRUE)
  ) %>%
  filter(count > 20, dest != 'HNL')
```


#### 9.2 Missing Values

Question: What happens if we don't set na.rm argument?

```{r}
flights %>% group_by(year,month,day) %>%
  summarize(mean = mean(dep_delay))
```

- Aggregation functions obey the usual rule of missing values.
..- If there's any missing vaue in the input, the output will be a missing value.
..- All aggregation functions have an na.rm argument, which removes the missing values

```{r}
flights %>% 
  group_by(year, month, day) %>%
  summarize(mean = mean(dep_delay, na.rm = TRUE))
```

- Missing values represent cancelled flights.

```{r}
not_cancelled <- flights %>% 
  filter(!is.na(dep_delay), !is.na(arr_delay))

not_cancelled %>%
  group_by(year, month, day) %>%
  summarize(mean = mean(dep_delay))
```


#### 9.3 Counts

- Let's look at the planes (identified by their tail number) that have the highest average delays.

```{r}
delays <- not_cancelled %>%
  group_by(tailnum) %>%
  summarize(
    delay = mean(arr_delay)
  )
```

```{r}
ggplot(data = delays, mapping = aes(x = delay)) +
  geom_freqpoly(binwidth = 10)
```

Observations:
- There are some planes that have an average delay of 5 hours (300 minutes)!

Let's draw a scatterplot of number of flights versus average delay.

```{r}
delays <- not_cancelled %>%
  group_by(tailnum) %>%
  summarize(
    delay = mean(arr_delay, na.rm = TRUE),
    n = n()
  )
```


```{r}
ggplot(data = delays, mapping = aes(x = n, y = delay)) +
  geom_point(alpha = 1/10)
```


#### 9.4 Useful Summary Functions

1. Measures of location

```{r}
# prepare not_cancelled flights
not_cancelled <- flights %>%
                    filter(!is.na(dep_delay), !is.na(arr_delay))
```

```{r}
not_cancelled %>%
  group_by(year,month,day) %>%
  summarize(
    avg_delay1 = mean(arr_delay),
    avg_delay2 = mean(arr_delay[arr_delay > 0 ])
  )
```

```{r}
flights$dep_time>0
```

```{r}
flights$dep_time[flights$dep_time>0]
```

```{r}
x <- c(1,2,3,4,5,6,7,8,9,10)
```

```{r}
x[x>4]
```

```{r}
x[3:6]
```

2. Measures of Spread sd(x), IQR(x), mad(x)

> Why is distance to some destinations more variable than to others?

```{r}
not_cancelled %>%
  group_by(dest) %>%
  summarize(distance_sd = sd(distance)) %>%
  arrange(desc(distance_sd))
```

3. Measures of rank min(x), quantile(x,0.25), max(x)

- Quantiles are a generalization of the median
- quantile(x,0.25) will find a value of x that is greater than 25% of the values and less than the remaining 75%

> When do the first and last flights leave each day ?

```{r}
not_cancelled %>%
  group_by(year,month,day) %>%
  summarize(
    first = min(dep_time),
    last = max(dep_time)
  )
```

4. Measures of position first(x), nth(x,2), last(x)

> find the first and last departure of each day

```{r}
not_cancelled %>%
  group_by(year,month,day) %>%
  summarize(
    first_dep = first(dep_time),
    last_dep = last(dep_time)
  )
```

> These functions are complementary to filtering on ranks.

```{r}
not_cancelled %>%
  group_by(year, month, day) %>%
  mutate(r = min_rank(desc(dep_time))) %>%
  filter(r %in% range(r))
```


5. Counts

```{r}
# which destinations have the most carriers?
not_cancelled %>%
  group_by(dest) %>%
  summarize(carriers = n_distinct(carrier)) %>%
  arrange(desc(carriers))
```

```{r}
# if you just need the count
not_cancelled %>%
  count(dest)
```

6. Counts and proportions of logical values sum(x > 10), mean(y == 0)

- When used with numeric functons, TRUE is converted to 1 and FALSE to 0
- This makes sum() and mean() very useful: sum(x) gives the number of TRUEs in x, and mean(x) gives the proportion

```{r}
# how many flights left before 5am? (these usually indicate delayed flights from the previous day)
not_cancelled %>%
  group_by(year, month, day) %>%
  summarize(n_early = sum(dep_time < 500))
```


```{r}
# What proportion of flights are delayed by more than a hour?
not_cancelled %>%
  group_by(year, month, day) %>%
  summarize(hour_perc = mean(arr_delay > 60, na.rm = TRUE))
```


#### 9.5 Grouping by Multiple Variables

```{r}
flights %>% group_by(year, month, day) %>% summarize(flights = n())
```

```{r}
# to find the number of flights per month
flights %>% group_by(year, month, day) %>% summarize(flights = n())
```

```{r}
flights %>% group_by(year, month, day) %>% summarize(flights = n()) %>% summarize(flights = sum(flights))
```


```{r}
flights %>% group_by(year, month, day) %>% summarize(flights = n()) %>% summarize(flights = sum(flights)) %>% summarize(flights = sum(flights))
```


```{r}
flights %>% group_by(year, month) %>% summarize(flights = n())
```


#### 9.6 Ungrouping

```{r}
flights %>%
  group_by(year, month, day) %>%
  summarize(count = n())
```


```{r}
flights %>%
  group_by(year, month, day) %>%
  ungroup() %>%
  summarize(count = n())
```

```{r}
flights %>%
  summarize(count = n())
```

### 10. Grouped Mutates (and Filters)

1. Find the worst members of each group - (year, month ,day) - greater arr_delays - 10 members

```{r}
flights %>% group_by(year,month,day) %>% mutate(rank = rank(desc(arr_delay))) %>% select(rank,arr_delay,everything()) %>% filter(rank<10)
```



```{r}
flights %>% 
  group_by(year,month,day) %>%
  filter(rank(desc(arr_delay))<10)
```
```{r}
flights %>%
  filter(month==1,day==1) %>%
  arrange(desc(arr_delay)) %>%
  head(n = 9)
```

```{r}
flights %>% 
  group_by(year,month,day) %>%
  arrange(desc(arr_delay)) %>%
  filter(month ==1, day ==1)
```

```{r}
flights %>% group_by(year,month,day) %>% arrange(year,month,day,desc(arr_delay)) %>% select(arr_delay, everything())
```

```{r}
flights %>% group_by(year,month,day) %>% mutate(rank = rank(desc(arr_delay))) %>% select(rank,arr_delay,everything()) %>% filter(rank<10)
```


2. Find all groups bigger than a threshold.

```{r}

```
this commands works


```{r}
flights%>%group_by(dest)%>%summarise(count=n())
```

```{r}
flights%>%group_by(dest)%>%filter(n()>365)%>%filter(dest=='ANC')
```
```{r}
flights%>%group_by(dest)%>%filter(n()>365)%>%filter(arr_delay>0)%>%mutate(pro=arr_delay/sum(arr_delay))%>%select(year,month,day,dest,arr_delay,pro)
```
```{r}
flights %>% 
  group_by(dest) %>%
  filter(n()>365) %>%
  filter(arr_delay>0) %>%
  summarize(sum = sum(arr_delay))
```
