---
title: "Unit II - Exercise"
author: "Suganya"
date: '2022-10-06'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(dplyr)
library(nycflights13)
```
```{r}
flights
```

1. Find all flights that: 
a. Had an arrival delay of two or more hours

```{r}
filter(flights, arr_delay >= 120)

```


B. Flew to Houston (IAH or HOU) 
```{r}
filter(flights, dest == "IAH" | dest == "HOU")
filter(flights, dest %in% c("IAH", "HOU"))

```



C. Were operated by United, American, or Delta
```{r}
filter(flights, carrier %in% c("AA", "DL", "UA"))

```

D. Departed in summer (July, August, and September)

```{r}
filter(flights, month >= 7, month <= 9)
filter(flights, month %in% 7:9)

```

E. Arrived more than two hours late, but didn’t leave late
```{r}
filter(flights, arr_delay > 120, dep_delay <= 0)

```


F. Were delayed by at least an hour, but made up over 30 minutes in flight
```{r}
filter(flights, dep_delay >= 60, dep_delay - arr_delay > 30)

```

G. Departed between midnight and 6 a.m. (inclusive)
```{r}
filter(flights, dep_time <= 600 | dep_time == 2400)


```

2. Another useful dplyr filtering helper is between(). What does it do? Can you use it to simplify the code needed to answer the previous challenges?
```{r}
filter(flights, between(month, 7, 9))

```


How many flights have a missing dep_time? What other variables are missing? What might these rows represent?
```{r}
filter(flights, is.na(dep_time))

```
Notably, the arrival time (arr_time) is also missing for these rows. These seem to be

cancelled flights


Why is NA ^ 0 not missing? Why is NA | TRUE not missing? Why is FALSE & NA not missing?

Can you figure out the general rule? (NA * 0 is a tricky counterexample!)


NA ^ 0 == 1 since for all numeric values x^0=1.

NA | TRUE is TRUE because the value of the missing TRUE or FALSE, x or TRUE is TRUE for

all values of x.

Likewise, anything and FALSE is always FALSE.

Since x∗0=0 for all finite, numeric x, So NA * 0 = 0

1. How could you use arrange() to sort all missing values to the start? (Hint: use is.na().) 
```{r}
arrange(flights, dep_time)
#Using desc() does not change that.
arrange(flights, desc(dep_time))
arrange(flights, desc(is.na(dep_time)), dep_time)
```

2. Sort flights to find the most delayed flights. Find the flights that left earliest.
```{r}
arrange(flights, desc(dep_delay))

arrange(flights, dep_delay)

```


3. Sort flights to find the fastest flights.
```{r}
mph = distance / air_time * 60
head(arrange(flights, desc(mph)))

```


4. Which flights traveled the longest? Which traveled the shortest?
```{r}
arrange(flights, desc(distance))

arrange(flights, distance)

```

```{r}
arrange(flights, desc(air_time))
arrange(flights, air_time)

```

1. Brainstorm as many ways as possible to select dep_time, dep_delay, arr_time, and arr_delay from flights.
```{r}
select(flights, dep_time, dep_delay, arr_time, arr_delay)
select(flights, "dep_time", "dep_delay", "arr_time", "arr_delay")
select(flights, 4, 6, 7, 9)
select(flights, one_of(c("dep_time", "dep_delay", "arr_time", "arr_delay")))
select(flights, starts_with("dep_"), starts_with("arr_"))
select(flights, ends_with("time"), ends_with("delay"))
select(flights, contains("_time"), contains("arr_"))


```

2. What happens if you include the name of a variable multiple times in a select() call?
```{r}
select(flights, year, month, day, year, year)

```

3. What does the one_of() function do? Why might it be helpful in conjunction with this vector?
```{r}
vars <- c("year", "month", "day", "dep_delay", "arr_delay") 
select(flights, one_of(vars))

```

4. Does the result of running the following code surprise you? How do the select helpers deal with case by default? How can you change that default?
```{r}
select(flights, contains("TIME"))
select(flights, contains("TIME", ignore.case = FALSE))

```

1. Currently dep_time and sched_dep_time are convenient to look at, but hard to compute

with because they’re not really continuous numbers. Convert them to a more convenient

representation of number of minutes since midnight.


1504 %/% 100     [1] 15   // Separate hours

1504 %% 100     [1] 4       // Separate Minutes

1504 %/% 100 * 60 + 1504 %% 100   //Convert  into total minutes

904    // Total minutes

904%% 1440   // Since midnight ,,,1440 Total minutes for midnight
```{r}
flights_times <- mutate(flights, dep_time_mins = (dep_time %/% 100 * 60 + dep_time %% 100) %% 1440, sched_dep_time_mins = (sched_dep_time %/% 100 * 60 + sched_dep_time %% 100) %% 1440 )
select( flights_times, dep_time, dep_time_mins, sched_dep_time, sched_dep_time_mins )

```

2. Compare air_time with arr_time - dep_time. What do you expect to see? What do you see? What do you need to do to fix it?
```{r}
flights_airtime <- mutate(flights, dep_time = (dep_time %/% 100 * 60 + dep_time %% 100) %% 1440, arr_time = (arr_time %/% 100 * 60 + arr_time %% 100) %% 1440, air_time_diff = arr_time - dep_time )
head(flights_airtime$air_time_diff )  ;   head(flights$air_time)

```

3. Compare dep_time, sched_dep_time, and dep_delay. How would you expect those three numbers to be related?
```{r}
flights_deptime <- mutate(flights, dep_time_min = (dep_time %/% 100 * 60 + dep_time %% 100) %% 1440, sched_dep_time_min = (sched_dep_time %/% 100 * 60 + sched_dep_time %% 100) %% 1440, dep_delay_diff =  dep_time_min - sched_dep_time_min )
head(flights_deptime$dep_delay_diff); head(flights$dep_delay) 

```


4. Find the 10 most delayed flights using a ranking function. How do you want to handle ties? Carefully read the documentation for min_rank()
```{r}
rankme <- tibble( x = c(10, 5, 1, 5, 5) )
rankme <- mutate(rankme, x_row_number = row_number(x), x_min_rank = min_rank(x), x_dense_rank = dense_rank(x) )
arrange(rankme, x)

```

```{r}
flights_delayed <- mutate(flights, dep_delay_min_rank = min_rank(desc(dep_delay))) 
flights_delayed <- filter( flights_delayed, !(dep_delay_min_rank > 10 ))
select(flights_delayed, dep_delay,dep_delay_min_rank)

```

5. What does 1:3 + 1:10 return? Why?
```{r}

```

6. What trigonometric functions does R provide?

All trigonometric functions are all described in a single help page, named Trig. 
R provides functions for the three primary trigonometric functions: sine (sin())), cosine (cos()), and tangent (tan()). 
The input angles to all these functions are in  radians.



1. Brainstorm at least five different ways to assess the typical delay characteristics

of a group of flights. Consider the following scenarios:

 A flight is 15 minutes early 50% of the time, and 15 minutes late 50% of the time.
 
 
A flight is always 10 minutes late. 

 A flight is 30 minutes early 50% of the time, and 30 minutes late 50% of the time. 
 
99% of the time a flight is on time. 1% of the time it’s 2 hours late. 

Which is more important: arrival delay or departure delay?


In many scenarios, arrival delay is more important. 

In most cases, being arriving late is more costly to the passenger since it could disrupt the next stages of their travel, such as connecting flights or scheduled meetings.

If a departure is delayed without affecting the arrival time, this delay will not  affects plans .



2. Come up with another approach that will give you the same output as 

i.not_cancelled %>% count(dest) and 

ii.not_cancel led %>% count(tailnum, wt = distance) 

(without using count())

```{r}
not_cancelled <- flights %>% filter(!is.na(dep_delay), !is.na(arr_delay))
not_cancelled %>% group_by(dest) %>% summarise(n = n())
not_cancelled %>% group_by(tailnum) %>% summarise(n = sum(distance))

```

3. Our definition of cancelled flights (is.na(dep_delay) | is.na(arr_delay) ) is slightly suboptimal. Why? Which is the most important column?

If the dep_delay is NULL then the flights are cancelled without any doubts.

But arr_delay is NULL then a flight may crashes, or if it is redirected and lands in an

airport other than its specified destination.

So the arr_delay column remains NULL for all diverted flights.



4. Look at the number of cancelled flights per day. Is there a pattern? Is the proportion of cancelled flights related to the average delay?
```{r}
cancelled_per_day <- flights %>% mutate(cancelled = (is.na(arr_delay) | is.na(dep_delay))) %>% 
group_by(year, month, day) %>% 
summarise( cancelled_num = sum(cancelled), flights_num = n() )

ggplot(cancelled_per_day) +
  geom_point(aes(x = flights_num, y = cancelled_num)) 

```


```{r}
cancelled_and_delays <-   flights %>%
  mutate(cancelled = (is.na(arr_delay) | is.na(dep_delay))) %>%
  group_by(year, month, day) %>%
  summarise(    cancelled_prop = mean(cancelled),
    avg_dep_delay = mean(dep_delay, na.rm = TRUE),
    avg_arr_delay = mean(arr_delay, na.rm = TRUE)    ) 
ggplot(cancelled_and_delays) +
  geom_point(aes(x = avg_dep_delay, y = cancelled_prop))
ggplot(cancelled_and_delays) +
  geom_point(aes(x = avg_arr_delay, y = cancelled_prop))

```

1. Refer back to the table of useful mutate and filtering functions. Describe how each

operation changes when you combine it with grouping. 
```{r}
tibble(x = 1:9, group = rep(c("a", "b", "c"), each = 3)) %>%  mutate(rnk = min_rank(x)) %>%  group_by(group) %>%  mutate(rnk2 = min_rank(x))
tibble(x = 1:9, group = rep(c("a", "b", "c"), each = 3)) %>% mutate(y = x + 2) %>% group_by(group) %>% mutate(z = x + 2)

```

2. What time of day should you fly if you want to avoid delays as much as possible?


The flight which scheduled  earlier , its expected delay become low. This is because  earlier flight’s delays will affect later flights. So group by an hour to know the earlier flights.

```{r}
flights %>% group_by(hour) %>% 
summarise(arr_delay = mean(arr_delay, na.rm = TRUE)) %>% 
arrange(arr_delay)

```

3. For each destination, compute the total minutes of delay. For each flight, compute the proportion of the total delay for its destination.
```{r}
flights %>% filter(arr_delay > 0) %>% group_by(dest) %>% 
mutate( arr_delay_total = sum(arr_delay), arr_delay_prop = arr_delay / arr_delay_total ) %>% 
select(dest, month, day, dep_time, carrier, flight, arr_delay, arr_delay_prop) %>% arrange(dest, desc(arr_delay_prop))

```

