---
title: 'Unit - I : Data Visualization with ggplot2'
author: 'Teacher : '
output: html_document
---

### UNIT-WISE LECTURE PRESENTATION RMD

#### Unit - 1 - DATA VIZUALIZATION WITH ggplot2

##### INTRODUCTION

  1. What is Data Science ?   
         
     Data science is an exciting discipline that allows you to turn raw data into understanding,insight and knowledge.
  
  2. What is R ? Rstudio?
     
     - R is a programming Language. 
     - It is Mainly used for Statistics Communication.
     - R studio is an IDE (Integrated Development Environment).
     
  3.What is TidyVerse ?   
     
     - Tidyverse is a library. 
     - It has collection of R packages designed for Data Science.
     
  4.How to Install a package ?

```{r eval=FALSE, include=FALSE, paged.print=FALSE}
install.packages("tidyverse")
```

  5.How to Attach a package to the current R session?

```{r include=FALSE}
library(tidyverse)
```

  6.How to Add dataset into R(mpg dataset)?

```{r}
library(tidyverse)
```
```{r}
data(mpg)
```


7. Data exploration of mpg dataset ?

```{r}
mpg
```

      a. View() - Invoke a spreadsheet-style data viewer on a matrix-like R object. 

```{r include=FALSE}
library(tidyverse)
data(mpg)
View(mpg)

```


     b. head() - returns the first 6 rows of the data frame
     
```{r}
head(mpg)

```


     c. tail() - returns the last 6 rows of the data frame
     
```{r}
tail(mpg)
```


     d. dim() - returns the dimensions of the data frame
     
```{r}
dim(mpg)
```


     e. colnames() - returns the column names of the data frame

```{r}
colnames(mpg)
```

     f. rownames() - returns the row names of the data frame
     
```{r}
rownames(mpg)
```


8.How to get the working directory in R? getwd()

```{r}
getwd()
```


9.How to set the working directory in R? setwd()

```{r}
#setwd("/home/sathish/r4ds")
```


10.ggplot() function

```{r}
ggplot(data=mpg) + 
  geom_point(mapping=aes(x=displ,y=hwy))
```
  
###### Add libraries

```{r warning=FALSE, include=FALSE}
library(tidyverse)
```


##### AESTHETIC MAPPINGS

###### Question: 

`do cars with big engines use more fuel than cars with small engines?`

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy))
```

**Observations:**

> as engine size increases, we see mileage decreasing
> it holds a negative relationship

###### Question : 

`In The above plot, one group of points(highlighted in red) seems to fall outside of the linear trend. These cars have a higher mileage than you might expect. How can you explain these cars?`

**aesthetic** An aesthetic is a visual property of the objects in your plot. viz.,size, shape and color.

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, color = class))
```


```{r message=FALSE, warning=FALSE}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, size = class))
```

```{r message=FALSE, warning=FALSE}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, shape = class))
```

**Observations: **

> there are only six shapes. 
> suv class have not been assigned any shape. 
> always by default, six shapes alone get assigned.


```{r message=FALSE, warning=FALSE}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, alpha = class))
```


```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy), color = 'red')
```

**Observations:**

> here the color doesn't convey information about a variable, but only changes the appearance of the plot. 
> it is not an aesthetic property of the graph.

##### GEOMETRIC OBJECTS

###### Question: How are these plots similar?

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy))
```

```{r}
ggplot(data = mpg) + geom_smooth(mapping = aes(x = displ, y = hwy))
```


##### geom_smooth() linetype argument

```{r}
count(mpg,drv)
```


```{r}
ggplot(data = mpg) + geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv))
```

```{r}
ggplot(data = mpg) + geom_smooth(mapping = aes(x = displ, y = hwy))

ggplot(data = mpg) + geom_smooth(mapping = aes(x = displ, y = hwy, group = drv))

ggplot(data = mpg) + geom_smooth(mapping = aes(x = displ, y = hwy, color = drv))

ggplot(data = mpg) + geom_smooth(mapping = aes(x = displ, y = hwy, color = drv), show.legend = FALSE)
```

- linetype argument : displays regression line for each category of the variable with a legend
- group argument : displays same line type for each of the category but different regression lines
- se argument: confidence interval
- geom_smooth() : function to add best fit regression line in the layer of ggplot()
- color: displays regression line with different color legends for each category of the variable

###### To display multiple geoms in the same plot

- ggplot(data = mpg) : created an empty canvas
- + geom_point() : created a scatterplot 
- + geom_smooth(): created a line (regression line)


```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ,y = hwy)) +
  geom_smooth(mapping = aes(x = displ, y = hwy))
```

###### to remove the confidence interval around the regression line

```{r}
ggplot(data = mpg) + geom_smooth(mapping = aes(x = displ,y = hwy), se = FALSE)
```

###### global mappings vs local mappings

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point() +
  geom_smooth()
```

###### global mappings + local mappings

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(mapping = aes(color = class)) +
  geom_smooth()
```

###### local data argument in geom_smooth()

```{r}
mpg_subcompact <- filter(mpg,class == "subcompact")

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(mapping = aes(color = class)) +
  geom_smooth(data = mpg_subcompact, se = FALSE)
```

```{r}
count(mpg,class)
```


###### to understand how filter works

- filter() function from dplyr package

```{r}
filter(mpg, class == "suv")
```


##### STATISTICAL TRANSFORMATIONS

###### diamonds dataset

```{r}
data(diamonds)
```

###### dataframe - diamonds dataset

```{r}
diamonds
```

**Observations on dataset diamonds:**

- data comprises of "prices of 50,000 round cut diamonds"
- variables are 

```{r}
colnames(diamonds)
```

  - has a dimension of 53940 x 10
  
```{r}
dim(diamonds)
```

  - min and max values of price variable in diamonds dataset
  
```{r}
range(diamonds$price)

min(diamonds$price)

max(diamonds$price)
```

###### to select a single variable from the entire dataset

```{r eval=FALSE, include=FALSE}
diamonds$color
```

  - a *geom* is the geometrical object that a plot uses to represent data.



```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut))
```

  - y variable - count ( statistical transformation )

```{r}
count(diamonds, cut)
```

```{r}
ggplot(data = diamonds) +
  stat_count(mapping = aes(x = cut))
```

###### Creating a new tibble using tribble() function
  
    - tribble() is used to create a new tibble for a smaller dataset or tables of data.
    - tibble is a data frame. It just a cleaner representation of a data frame.
    - tibble is different from a traditional data frame in two things (printing and subsetting)

```{r}
demo <- tribble(
  ~a, ~b,
  "bar_1", 20,
  "bar_2",30,
  "bar_3",40
)
```

```{r}
demo
```

```{r}
ggplot(data = demo) +
  geom_bar(mapping = aes(x = a))
```

```{r}
count(demo,a)
```


```{r}
ggplot(data = demo) +
  geom_bar(mapping = aes(x = a,y = b), stat = "identity")
```

 - stat argument with value "identity" helps us to map the height of the bars  with the raw values of a y variable.
 
###### overriding the default mapping from transformed variables to aesthetics

  - changing group values doesn't make any changes.
  - but removing group = 1 makes the bars to have same heights (prop = 1.0)

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, y = ..prop.., group = 1))
```

###### stat_summary()

```{r}
#ggplot(data = diamonds) +
  #stat_summary(mapping = aes(x = cut, y = depth),
               #fun.ymin = min,
               #fun.ymax = max,
               #fun.y = median())
```

**Some of the stat functions:**

```{r}
range(diamonds$depth)
min(diamonds$depth)
max(diamonds$depth)
median(diamonds$depth)
```

```{r}
diamonds %>% group_by(cut) %>% summarize(min = min(depth),max = max(depth),med = median(depth))
```

```{r}
good <- diamonds %>% filter(cut == "Good")
```


##### POSITION ADJUSTMENTS

###### Color a bar chart using either the color aesthetic, or fill

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, color = cut))

ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = cut))
```

###### Stacked bars

  - What happens when we map the fill aesthetic to another variable , like clarity
  
```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = clarity))
```

```{r}
diamonds %>% filter(cut == "Ideal" & clarity == "IF")
```


###### Non-Stacked Bars

  - when we don't need stacked bars, we can use one of the three options
    
        1. position = "identity" -> will place each object exactly where it falls in the context of the graph.
        
```{r}
ggplot(data = diamonds, mapping = aes(x = cut, fill = clarity)) +
  geom_bar(alpha = 1/5)

ggplot(data = diamonds, mapping = aes(x = cut, fill = clarity)) +
  geom_bar(alpha = 1/5, position = "identity")
```
        
  - slightly transparent = setting alpha to a small value or
  
```{r}
ggplot(data = diamonds, mapping = aes(x = cut, fill = clarity)) +
  geom_bar(alpha = 1/5, position = "identity")
```
  
  - completely transparent = setting fill = NA
  
```{r}
ggplot(data = diamonds, mapping = aes(x = cut, color = clarity)) +
  geom_bar(fill = NA, position = "identity")
```
  
        
        2. position = "fill" -> works like stacking, but makes each set of stacked bars the same height
        
```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "fill")
```
        
        
        3. position = "dodge" -> places overlapping objects directly beside one another.
                              -> easier to compare individual values

```{r}

ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "dodge")

```


###### The Problem of *Overplotting*

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy))
```

  - plot displays only 126 points, even though there are 234 observations in the dataset
  - when points overlap each other, it is called "overplotting"
  
  - inorder to solve the overplotting problem, we can set the position argument with the value "jitter" which adds some random noise to each data point.
  
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy), position = "jitter")
```
  
```{r}
ggplot(data = mpg) +
  geom_jitter(mapping = aes(x = displ, y = hwy))
```


##### COORDINATE SYSTEMS

1. coord_flip() switches the x-axes and y-axes

```{r}
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) +
  geom_boxplot()
```

```{r}
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) +
  geom_boxplot() +
  coord_flip()
```

2. coord_quickmap() -> sets the aspect ratio correctly for maps

```{r}

```


3. coord_polar() -> uses polar coordinates. It is most commonly used in pie charts.

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = cut), show.legend = FALSE, width = 1)+
  theme(aspect.ratio = 1) +
  labs(x = NULL, y = NULL) + 
  coord_flip()

ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = cut), show.legend = TRUE, width = 1)+
  theme(aspect.ratio = 1) +
  labs(x = NULL, y = NULL) +
  coord_flip() +
  coord_polar(theta = "x")

```


```{r}
count(diamonds,cut)
```


```{r}
ggplot(data = diamonds) + geom_bar(mapping = aes(x = cut, fill = cut), width = 1) + coord_polar()
```


